### 2015-06-24 manuscript/coverletter.tex
Updated coverletter, results in main text

\documentclass[12pt]{letter}

\usepackage[britdate,alyssa-signature]{canterbury-letter}
\usepackage{times}
\usepackage{letterbib}
\usepackage{geometry}
\usepackage[round]{natbib}
\usepackage{graphicx}
\geometry{a4paper}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{authblk}
\usepackage[running]{lineno}
\usepackage{amsmath,amsfonts,amssymb}
% \usepackage[margin=10pt,font=small,labelfont=bf]{caption}

%\usepackage{natbib}
% \bibpunct[; ]{(}{)}{;}{a}{,}{;}

\newenvironment{refquote}{\bigskip \begin{it}}{\end{it}\smallskip}

\newenvironment{figure}{}

\position{PhD Candidate}
\department{School of Biological Sciences}
\location{Private Bag 4800}
\telephone{+64 3 364 2729}
\fax{+64 3 364 2590}
\email{alyssa.cirtwill@pg.canterbury.ac.nz}
\url{http://stoufferlab.org}
\name{Alyssa R. Cirtwill}

% \position{PhD student}
% \department{School of Biological Sciences}
% \location{Private Bag 4800}
% \telephone{+64 3 364 2729}
% \fax{+64 3 364 2590}
% \email{alyssa.cirtwill@pg.canterbury.ac.nz}
% \url{http://stoufferlab.org}
% \name{Ms. Alyssa R. Cirtwill}

\newcommand{\mytitle}{\emph{Generality in food webs scales with species richness but not with latitude}}
\newcommand{\myjournal}{\emph{Proceedings of the Royal Society B}}

\begin{document}

\begin{letter}{\bf Professor Spencer C.H. Barrett\\
               Editor-in-Chief\\
               Proceedings of the Royal Society B\\
               6-9 Carlton House Terrace\\
               London, UK\\
               SW1Y 5AG\\
                }

\opening{Dear Prof. Barrett:}

We are happy to invite you to consider our manuscript 
``\mytitle'' for publication in the Population and Community Ecology section of the \emph{\myjournal}.

Our study aims to combine two well-established but disparate ecological
themes, namely the analysis of macroecological gradients and the search for
general laws governing the structure of food webs (networks of who eats whom
within a community). In particular, we were interested in the possibility of a
link between the well-known latitudinal gradients in species richness of some
taxa (e.g., birds, land plants) and the scaling relationships between species
richness and food-web structure. With regard to the former, one  hypothesized
mechanism for greater species richness in the tropics is the latitude-niche-
breadth hypothesis, which predicts that the greater stability of temperature
and productivity in the tropics will permit species to evolve narrower niches
and therefore more finely partition habitats. If this hypothesis holds true,
then tropical species should be more specialized than their high-latitude
counterparts. In terms of a food web, this means that we would naively expect
that tropical species will have relatively fewer feeding links than polar
species. However, the number of feeding links in which a species is involved
is known to scale with the number of species in the community food web
following a power law. Therefore we also naively expect that species in
tropical food webs will participate in many feeding links by virtue of their
speciose communities. This apparent paradox can be resolved by adding an
effect of latitude to the exponent of the power law. We tested for the impact
of such an exponent, as well as potential interactions between the effects of
latitude and ecosystem type on scaling laws.

In general we found little evidence of an effect of latitude on scaling
relationships between diet breadth and species richness, implying that
latitudinal variation in food-web structure may be driven purely by
latitudinal gradients in species richness. If diet breadth may be taken as a
measure of a species' Eltonian (biotic) niche, this means that our results are
inconsistent with the latitude-niche-breadth hypothesis. Ecosystem type, on
the other hand, had a greater impact on scaling exponents, suggesting that
different processes may structure food webs in different habitats. Indeed,
freshwater food webs departed from the overall trend and showed strong effects
of latitude on scaling relationships. In these food webs alone was the
latitude-niche-breadth hypothesis supported by our results.

By linking macroecology and food-web ecology, we believe that our study will
be of interest to ecologists, theory-focused biologists, and to general
readers of \emph{\myjournal}. Our results are potentially applicable to
researchers exploring community structure and specialization in a variety of
different systems and also provide a new perspective on the debate over the
latitude-niche-breadth hypothesis. We hope you will agree that this broad
relevance and the novelty of our results justify consideration of our
manuscript's inclusion in \emph{\myjournal}.

Thank you again for your consideration.

\closing{Regards,}

\end{letter}

% \newpage

% \setcounter{page}{1}

\end{document}