### 2015-07-01 manuscript/bigtable.tex
Big big table

% Table generated by Excel2LaTeX from sheet 'Sheet1'
\begin{table}[htbp]
  \centering
  \caption{Add caption}
    \begin{tabular}{rrrrrrrrr}
    \toprule
    Web ID & Collection & Original Source & Location & Ecosystem & Latitude(degrees) & Used  & Reason for rejection & Notes \\
    \midrule
    Alamitos & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Harrison1995} & USA   & stream & 37,12,25 N & Yes   &       &  \\
    Alford & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Havens1992} & USA   & lake  & 44,15,41.59 N & Yes   &       &  \\
    Beaver & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Havens1992} & USA   & lake  & 44,18,58.13 N & Yes   &       &  \\
    Bere  & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Woodward2008} & England & stream & 50,45,7 N & Yes   &       &  \\
    Bridge & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Havens1992} & USA   & lake  & 44,09,42.66 N & Yes   &       &  \\
    Broad & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Townsend1998} & New Zealand & stream & 45.896938 S & Yes   &       &  \\
    Broadstone & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Woodward2005} & stream & stream & 51,01,05 N & Yes   &       &  \\
    Calero & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Harrison1995} & USA   & stream & 37,11,09 - 20 N & Yes   &       &  \\
    Canton & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Townsend1998} & New Zealand & stream & 45.752 S & No    & Duplicated in GlobalWeb (web 235) &  \\
    Carpinteria2006 & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Lafferty2006} & USA   & estuary & 34.41 - 34.40 N & Yes   &       &  \\
    Chub  & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Havens1992} & USA   & lake  & 43,31,39.36 N & Yes   &       &  \\
    Coachella & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Polis1991a} & desert & terrestrial & 33,54 N & Yes   &       &  \\
    Connery & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Havens1992} & USA   & lake  & 44,18,42.10 N & Yes   &       &  \\
    Corte & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Harrison1995} & USA   & stream & 37,56,28 - 37,57,39 N & Yes   &       &  \\
    Coyote & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Harrison1995} & USA   & stream & 37,53,11.66 - 37,53,17.06 N & Yes   &       &  \\
    German & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Townsend1998} & New Zealand & stream & 44.97 S & No    & Duplicated in GlobalWeb (web 227) &  \\
    GrandeCl & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{CattinBlandenier2004} & wetland & terrestrial & 46,47,09 N & Yes   &       &  \\
    GrandeSc & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{CattinBlandenier2004} & wetland & terrestrial & 46,47,09 N & Yes   &       &  \\
    Guadalupe & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Harrison1995} & USA   & stream & 37,27,54 - 37,21,17 N & Yes   &       &  \\
    Hainich   & ~\textbackslash{}citet{Riede2011} & Unpublished & Germany & terrestrial &       & No    & Could not locate original source &  \\
    LittleKyeBurn & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Townsend1998} & New Zealand & stream & 45 S  & No    & Duplicated in GlobalWeb (web 230) &  \\
    LittleRock & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Martinez1991} & USA   & lake  & 45,44,07.94 N & Yes   &       &  \\
    Mondego & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Patricio2006} & Portugal & estuary & 40,08 N & Yes   &       &  \\
    Montane Forest, Arizona & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Rasmussen1941} & USA   & terrestrial & 36,5-37 N & No    & Duplicated in GlobalWeb (web60) &  \\
    Reef  & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Opitz1996} & Caribbean & marine & 18,29 - 17,53 N & Yes   &       &  \\
    Sierra & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Brose2005} & USA   & lake  & 36,50 - 37,02 N & Yes   &       &  \\
    SimberloffE1 & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Simberloff1976} & Mangrove islands & terrestrial & 24,40,37.29 N & Yes   &       &  \\
    SimberloffE2 & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Simberloff1976} & Mangrove islands & terrestrial & 24,40,33 N & Yes   &       &  \\
    SimberloffE3 & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Simberloff1976} & Mangrove islands & terrestrial & 24,41,23.82 N & Yes   &       &  \\
    SimberloffE7 & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Simberloff1976} & Mangrove islands & terrestrial & 25,15 N & Yes   &       &  \\
    SimberloffE9 & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Simberloff1976} & Mangrove islands & terrestrial & 25,01,59.64 - 25,02,55.44 N & Yes   &       &  \\
    SimberloffST2 & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Simberloff1976} & Mangrove islands & terrestrial & 24,36,16 N & Yes   &       &  \\
    Skipwith & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Warren1989} & England & lake  & 53,40 N & Yes   &       &  \\
    Stink & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Havens1992} & USA   & lake  & 43,37,55.88 N & Yes   &       &  \\
    StMarks & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Christian1999} & USA   & estuary & 30,05 N & No    & Duplicated in GlobalWeb (web 288) &  \\
    Stony  & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Townsend1998} & New Zealand & stream & 45.58 S & No    & Duplicated in GlobalWeb (web 231) &  \\
    TreleaseWoods & ~\textbackslash{}citet{Riede2011} & \multicolumn{1}{l}{~\textbackslash{}citet{Twomey1945}} & USA   & terrestrial & 40,09,47 N & No    & Duplicated in GlobalWeb (web 59) &  \\
    Tuesday & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Jonsson2005} & USA   & lake  & 46,13 N & Yes   &       &  \\
    Weddell Sea & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Brose2005} & Antarctica & marine & 58,20.8 - 60,10.4 S & No    & Duplicated in GlobalWeb (web 341) &  \\
    Ythan & ~\textbackslash{}citet{Riede2011} & ~\textbackslash{}citet{Huxham1996} & UK    & marine &       & No    & Duplicated in Dunne (YthanDunne) &  \\
    BahiaDunne & ~\textbackslash{}citet{Dunne2013} & ~\textbackslash{}citet{Hechinger2011} & Mexico & estuary & 30.43 - 30.46 N & Yes   &       & Parasites removed \\
    CarpinteriaDunne & ~\textbackslash{}citet{Dunne2013} & ~\textbackslash{}citet{Hechinger2011} & USA   & estuary & 34.41 - 34.40 N & Yes   &       & Parasites removed \\
    EsteroDunne & ~\textbackslash{}citet{Dunne2013} & ~\textbackslash{}citet{Hechinger2011} & Mexico & estuary & 31.78 - 31.69 N & Yes   &       & Parasites removed \\
    FjordDunne & ~\textbackslash{}citet{Dunne2013} & ~\textbackslash{}citet{Zander2011} & Germany & estuary & 54.92 - 54.75 N & Yes   &       & Parasites removed \\
    OtagoDunne & ~\textbackslash{}citet{Dunne2013} & ~\textbackslash{}citet{Mouritsen2011} & New Zealand & estuary & 45.78 - 45.88 S & Yes   &       & Parasites removed \\
    SyltDunne & ~\textbackslash{}citet{Dunne2013} & ~\textbackslash{}citet{Thieltges2011} & Denmark & estuary & 54.83 - 55.12 N & Yes   &       & Parasites removed \\
    YthanDunne & ~\textbackslash{}citet{Dunne2013} & ~\textbackslash{}citet{Huxham1996} & Scotland & estuary & 57,18,32 N & Yes   &       & Parasites removed \\
    1     & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Qazim1970}} & India & \multicolumn{1}{l}{estuary} &       & No    & Could not locate original source &  \\
    2     & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Day1967}} & South Africa & \multicolumn{1}{l}{estuary} &       & No    & Could not locate original source &  \\
    3     & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Woodwell1967}} & USA   & \multicolumn{1}{l}{estuary} & 41,03,11 N & Yes   &       &  \\
    4     & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Johnston1956}} & USA   & \multicolumn{1}{l}{estuary} &       & No    & Sink web &  \\
    5     & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Teal1962}} & USA   & \multicolumn{1}{l}{estuary} & 31,28,38 N & Yes   &       &  \\
    6     & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{MacGinitie1935}} & USA   & \multicolumn{1}{l}{estuary} & 36,48,18 N & Yes   &       &  \\
    7     & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Kremer1978}} & USA   & \multicolumn{1}{l}{marine} & 41,45 - 41,30 N & Yes   &       &  \\
    8     & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Nixon1973}} & USA   & \multicolumn{1}{l}{estuary} & 41,33,00 N & Yes   &       &  \\
    9     & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Kitching1967}} & Ireland & \multicolumn{1}{l}{stream} &       & No    & Could not locate original source &  \\
    10    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Menge1976}} & USA   & \multicolumn{1}{l}{marine} & 41 - 47 N & Yes   &       &  \\
    11    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Menge1976}} & USA   & \multicolumn{1}{l}{marine} & 41 - 47 N & Yes   &       &  \\
    12    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Menge1976}} & USA   & \multicolumn{1}{l}{marine} & 45,45 - 49 N & Yes   &       &  \\
    13    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Menge1976}} & USA   & \multicolumn{1}{l}{marine} & 45,45 - 49 N & Yes   &       &  \\
    14    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Walsh1967}} & Hawaii & \multicolumn{1}{l}{estuary} & 21,26,09 N & No    & Could not locate original source &  \\
    15    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Walsh1967}} & Hawaii & \multicolumn{1}{l}{estuary} & 21,26,09 N & No    & Could not locate original source &  \\
    16    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Copeland1974}} & USA   & \multicolumn{1}{l}{estuary} &       & No    & Could not locate original source &  \\
    17    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Hiatt1960}} & Marshall Islands - Pacific Ocean & \multicolumn{1}{l}{marine} & 4,30 - 15 N & No    & Generalized scheme &  \\
    18    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Niering1963a}} & Polynesia & \multicolumn{1}{l}{terrestrial} & 1,01,29 - 1,06,23 N & Yes   &       &  \\
    19    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Brooks1963}} & USA   & \multicolumn{1}{l}{lake} &       & No    & Could not locate original source &  \\
    20    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Knox1970}} & Antarctica & \multicolumn{1}{l}{terrestrial} &       & No    & Could not locate original source &  \\
    21    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Patten1979}} & Antarctica & \multicolumn{1}{l}{marine} & 70-85 S & Yes   &       &  \\
    22    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Summerhayes1923}} & Norway & \multicolumn{1}{l}{terrestrial} & 74,27,54 N & Yes   &       &  \\
    23    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Bird1930}} & Canada & \multicolumn{1}{l}{terrestrial} & 50,25,17 N & Yes   &       &  \\
    24    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Bird1930}} & Canada & \multicolumn{1}{l}{terrestrial} & 50,25,17 N & Yes   &       &  \\
    25    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Bird1930}} & Canada & \multicolumn{1}{l}{terrestrial} & 50,25,17 N & Yes   &       &  \\
    26    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Bird1930}} & Canada & \multicolumn{1}{l}{terrestrial} & 50,25,17 N & Yes   &       &  \\
    27    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Varley1970}} & England & \multicolumn{1}{l}{terrestrial} &       & No    & Could not locate original source &  \\
    28    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Paviour-Smith1956}} & New Zealand & \multicolumn{1}{l}{estuary} & 45,51,30 S & Yes   &       &  \\
    29    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Dunbar1953}} & Arctic seas & \multicolumn{1}{l}{marine} &       & No    & Generalized scheme &  \\
    30    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Mackintosh1964}} & Antarctica & \multicolumn{1}{l}{marine} &       & No    & Generalized scheme &  \\
    31    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Petipa1970}} & Black Sea & \multicolumn{1}{l}{marine} &       & No    & Could not locate original source &  \\
    32    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Petipa1970}} & Black Sea & \multicolumn{1}{l}{marine} &       & No    & Could not locate original source &  \\
    33    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Fryer1959}} & Malawi & \multicolumn{1}{l}{lake} & 11,36,39 S & Yes   &       &  \\
    34    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Jones1949}} & UK    & \multicolumn{1}{l}{stream} & 51,41 - 51,55 N & Yes   &       &  \\
    35    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Minshall1967}} & USA   & \multicolumn{1}{l}{stream} & 37,18,26 N & Yes   &       &  \\
    36    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Walsh1967}} & Hawaii & \multicolumn{1}{l}{estuary} & 21,26,09 N & No    & Could not locate original source &  \\
    37    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Clarke1967}} & USA   & \multicolumn{1}{l}{marine} & 41,02,49 N & Yes   &       &  \\
    38    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Fryer1959}} & Malawi & \multicolumn{1}{l}{lake} & 11,36,39 S & Yes   &       &  \\
    39    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Fryer1959}} & Malawi & \multicolumn{1}{l}{lake} & 11,36,39 S & Yes   &       &  \\
    40    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Harrison1962}} & Malaysia & \multicolumn{1}{l}{terrestrial} &       & No    & Generalized scheme &  \\
    41    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Parin1970}} & Israel & \multicolumn{1}{l}{marine} &       & No    & Could not locate original source &  \\
    42    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Vinogradov1978}} & Pacific Ocean & \multicolumn{1}{l}{marine} &       & No    & Study site unclear &  \\
    43    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Rosenthal1974}} & USA   & \multicolumn{1}{l}{marine} & 32,57 N & Yes   &       &  \\
    44    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Yanez1978}} & Mexico & \multicolumn{1}{l}{marine} & 17,26,21 N & Yes   &       &  \\
    45    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Tilly1968}} & USA   & \multicolumn{1}{l}{stream} & 41,22,55 N & Yes   &       &  \\
    46    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Patten1975}} & USA   & \multicolumn{1}{l}{lake} & 33,54 N & Yes   &       &  \\
    47    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Harris1980}} & USA   & \multicolumn{1}{l}{lake} &       & No    & Study site unclear &  \\
    48    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Simenstad1978}} & USA   & \multicolumn{1}{l}{marine} &       & No    & Extinct species &  \\
    49    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Simenstad1978}} & USA   & \multicolumn{1}{l}{marine} &       & No    & Extinct species &  \\
    50    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Nybakken1982}} & USA   & \multicolumn{1}{l}{marine} &       & No    & Generalized scheme &  \\
    51    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Dexter1947}} & USA   & \multicolumn{1}{l}{marine} & 42,38,56 N & Yes   &       &  \\
    52    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Paine1980a}} & USA   & \multicolumn{1}{l}{marine} & 58,27 N & Yes   &       &  \\
    53    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Paine1980a}} & USA   & \multicolumn{1}{l}{marine} & 48,19 - 48,37 N & Yes   &       &  \\
    54    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{Unpublished} & Barbados & \multicolumn{1}{l}{marine} &       & No    & Unpublished &  \\
    55    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Milne1972}} & Scotland & \multicolumn{1}{l}{estuary} &       & No    & Could not locate original source &  \\
    56    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Milne1972}} & Scotland & \multicolumn{1}{l}{estuary} &       & No    & Could not locate original source &  \\
    57    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Yanez1978}} & Mexico & \multicolumn{1}{l}{marine} & 17,26,21 N & Yes   &       &  \\
    58    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cohen1990}} & Russia & \multicolumn{1}{l}{lake} & 51,46 N & Yes   &       &  \\
    59    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Twomey1945}} & USA   & \multicolumn{1}{l}{terrestrial} & 40,09,47 N & Yes   &       &  \\
    60    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Rasmussen1941}} & USA   & \multicolumn{1}{l}{terrestrial} & 36,5-37 N & Yes   &       &  \\
    61    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Summerhayes1928}} & Norway & \multicolumn{1}{l}{terrestrial} & 77,52,30 N & Yes   &       &  \\
    62    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Summerhayes1928}} & Norway & \multicolumn{1}{l}{terrestrial} & 77,52,30 N & Yes   &       &  \\
    63    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Jones1950}} & UK    & \multicolumn{1}{l}{stream} & 52,25,08 N & Yes   &       &  \\
    64    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cummins1966}} & USA   & \multicolumn{1}{l}{stream} & 41,39,33 N & Yes   &       &  \\
    65    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Tsuda1972}} & Japan & \multicolumn{1}{l}{stream} & 34,21,20 - 34,23,21 N & Yes   &       &  \\
    66    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Mann1972a}} & England & \multicolumn{1}{l}{stream} & 51,30 N & Yes   &       &  \\
    67    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Carlson1968}} & USA   & \multicolumn{1}{l}{river} & 40,24,16 N & Yes   &       &  \\
    68    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Morgan1972}} & Scotland & \multicolumn{1}{l}{lake} &       & No    & Could not locate original source &  \\
    69    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cohen1990}} & Portugal & \multicolumn{1}{l}{estuary} &       & No    & Could not locate original source &  \\
    70    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Kemp1977}} & USA   & \multicolumn{1}{l}{estuary} & 28,54,08 N & Yes   &       &  \\
    71    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Sorokin1972}} & Russia & \multicolumn{1}{l}{lake} & 48,00 N & Yes   &       &  \\
    72    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Baril1983}} & Canada & \multicolumn{1}{l}{lake} & 46,02 N & Yes   &       &  \\
    73    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Schiemer1979}} & Austria & \multicolumn{1}{l}{lake} &       & No    & Could not locate original source &  \\
    74    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cohen1990}} & South Africa & \multicolumn{1}{l}{marine} & 34,15,00 S & Yes   &       &  \\
    75    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Pechlaner1972}} & Austria & \multicolumn{1}{l}{lake} & 47,11 N & Yes   &       &  \\
    76    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Schiemer1979}} & Austria & \multicolumn{1}{l}{lake} & 47,42 - 47,54 N & Yes   &       &  \\
    77    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cohen1990}} & Ethiopia & \multicolumn{1}{l}{lake} &       & No    & Could not locate original source &  \\
    78    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Burgis1972}} & Uganda & \multicolumn{1}{l}{lake} & 0     & Yes   &       &  \\
    79    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Sarvala1974}} & Finland & \multicolumn{1}{l}{lake} &       & No    & Could not locate original source &  \\
    80    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Sarvala1974}} & Finland & \multicolumn{1}{l}{lake} &       & No    & Could not locate original source &  \\
    81    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Hatanaka1977}} & Japan & \multicolumn{1}{l}{marine} &       & No    & Could not locate original source &  \\
    82    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cohen1990}} & France & \multicolumn{1}{l}{marine} &       & No    & Could not locate original source &  \\
    83    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cohen1990}} & Finland & \multicolumn{1}{l}{lake} &       & No    & Could not locate original source &  \\
    84    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Wilbur1972}} & USA   & \multicolumn{1}{l}{lake} & 42,28,42 N & Yes   &       &  \\
    85    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Mizuno1982}} & Malaysia & \multicolumn{1}{l}{lake} &       & No    & Could not locate original source &  \\
    86    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Hogetsu1979}} & Japan & \multicolumn{1}{l}{marine} & 34,36 - 35,38 N & Yes   &       &  \\
    87    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Bradstreet1982}} & Canada & \multicolumn{1}{l}{terrestrial} & 72,46,34.95 N & Yes   &       &  \\
    88    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Kuusela1980}} & Finland & \multicolumn{1}{l}{stream} &       & No    & Could not locate original source &  \\
    89    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Hartley1948}} & England & \multicolumn{1}{l}{stream} & 52,13 N & Yes   &       &  \\
    91    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cohen1990}} & Japan & terrestrial &       & No    & Could not locate original source &  \\
    92    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Swan1961}} & Tibet & terrestrial &       & No    & Generalized scheme &  \\
    93    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Pattie1966}} & USA   & terrestrial & 45 N  & Yes   &       &  \\
    94    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Brown1971}} & USA   & terrestrial &       & No    & Could not locate original source &  \\
    95    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Brown1975}} & USA   & terrestrial & 70,11,14.18 N & Yes   &       &  \\
    96    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cohen1990}} & Russia & terrestrial &       & No    & Could not locate original source &  \\
    97    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cohen1990}} & Russia & terrestrial &       & No    & Could not locate original source &  \\
    98    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Holm1980}} & Namibia & terrestrial &       & No    & Could not locate original source &  \\
    99    & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Howes1954}} & USA   & terrestrial & 32,36 N & Yes   &       &  \\
    100   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cohen1990}} & India & terrestrial &       & No    & Could not locate original source &  \\
    101   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cohen1990}} & \multicolumn{1}{l}{France} & lake  &       & No    & Could not locate original source &  \\
    102   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Shushkina1973}} & \multicolumn{1}{l}{Pacific Ocean} & marine & 10 S - 15 N & Yes   &       &  \\
    103   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Petipa1979}} & \multicolumn{1}{l}{Pacific Ocean} & marine &       & No    & Generalized scheme &  \\
    104   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Menge1976}} & \multicolumn{1}{l}{Panama} & marine & 4,48,04 N & Yes   &       &  \\
    105   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Edwards1982}} & \multicolumn{1}{l}{USA} & marine & 42,40 N & Yes   &       &  \\
    106   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Glynn1965}} & \multicolumn{1}{l}{USA} & marine &       & No    & Could not locate original source &  \\
    107   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Peterson1979}} & \multicolumn{1}{l}{USA} & marine & 39,52,02 N & Yes   &       &  \\
    108   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Hewatt1937}} & \multicolumn{1}{l}{USA} & marine & 36,37 N & Yes   &       &  \\
    109   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Castilla1981}} & \multicolumn{1}{l}{Chile} & marine &       & No    & Could not locate original source &  \\
    110   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Dexter1947}} & \multicolumn{1}{l}{USA} & marine & 42,38,56 N & Yes   &       &  \\
    111   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Dexter1947}} & \multicolumn{1}{l}{USA} & marine & 42,38,56 N & Yes   &       &  \\
    112   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Dexter1947}} & \multicolumn{1}{l}{USA} & marine & 42,38,56 N & Yes   &       &  \\
    113   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Dexter1947}} & \multicolumn{1}{l}{USA} & marine & 42,38,56 N & Yes   &       &  \\
    114   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Marshall1982}} & \multicolumn{1}{l}{Zimbabwe} & lake  & 17,54,11.90 S & Yes   &       &  \\
    115   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Koepcke1952}} & \multicolumn{1}{l}{Peru} & marine &       & No    & Could not locate original source &  \\
    116   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Hurlbert1972}} & \multicolumn{1}{l}{U.S.A.} & lake  & 35,09,19 S & Yes   &       &  \\
    117   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Zaret1973}} & \multicolumn{1}{l}{Panama} & lake  & 9,11,31.50 N & Yes   &       &  \\
    118   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Bindloss1972}} & \multicolumn{1}{l}{UK} & lake  & 56,41,18.71 N & Yes   &       &  \\
    119   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Mann1972}} & \multicolumn{1}{l}{UK} & stream & 51,27 N & Yes   &       &  \\
    120   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Moriarty1973}} & \multicolumn{1}{l}{Uganda} & lake  & 0,00,58 N & Yes   &       &  \\
    121   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{VanEs1977}} & \multicolumn{1}{l}{Germany} & estuary & 53,20 - 53,15 N & Yes   &       &  \\
    122   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Nixon1973}} & \multicolumn{1}{l}{USA} & estuary & 41,33,04 N & Yes   &       &  \\
    123   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Harris1972}} & \multicolumn{1}{l}{USA} & terrestrial & 40,45,28 N & Yes   &       &  \\
    124   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Collins1976}} & \multicolumn{1}{l}{USA} & stream & 44,34 N & Yes   &       &  \\
    125   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Landry1977}} & \multicolumn{1}{l}{Unspecified} & marine &       & No    & Generalized scheme &  \\
    126   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Larsson1978}} & \multicolumn{1}{l}{Norway} & lake  & 61,26,00 N & Yes   &       &  \\
    127   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Beaver1979}} & \multicolumn{1}{l}{Malaysia} & lake  &       & No    & Could not locate original source &  \\
    128   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Beaver1979}} & \multicolumn{1}{l}{Malaysia} & lake  &       & No    & Could not locate original source &  \\
    129   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{Unpublished} & \multicolumn{1}{l}{Unspecified} & terrestrial &       & No    & Unpublished &  \\
    130   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Odum1975}} & \multicolumn{1}{l}{USA} & estuary & 27,12 N & Yes   &       &  \\
    131   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Beaver1985}} & \multicolumn{1}{l}{Malaysia} & lake  & 4,12,37 N & Yes   &       &  \\
    132   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Beaver1985}} & \multicolumn{1}{l}{Malaysia} & lake  & 4,12,37 N & Yes   &       &  \\
    133   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Beaver1985}} & \multicolumn{1}{l}{Sri Lanka} & lake  & 7,52,22 N & Yes   &       &  \\
    134   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Beaver1985}} & \multicolumn{1}{l}{Madagascar} & lake  & 19 S  & Yes   &       &  \\
    135   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Beaver1985}} & Seychelles & lake  & 4,40,46 S & Yes   &       &  \\
    136   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Bradshaw1983}} & USA   & lake  &       & No    & Could not locate original source &  \\
    137   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Corker1984}} & Hong Kong & lake  & 22,14-22,17 N & Yes   &       &  \\
    138   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Kitching1983}} & UK    & lake  &       & No    & Could not locate original source &  \\
    139   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Kitching1983}} & Australia & lake  &       & No    & Could not locate original source &  \\
    140   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Siefert1979}} & Venezuela & terrestrial & 10,21 N & Yes   &       &  \\
    141   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Seifert1976}} & Costa Rica & terrestrial &       & No    & Competition network &  \\
    142   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Seifert1976}} & Costa Rica & terrestrial &       & No    & Competition network &  \\
    143   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Snow1958}} & USA   & lake  & 34,41,45 N & Yes   &       &  \\
    144   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Snow1958}} & Germany & lake  &       & No    & Could not locate original source &  \\
    145   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Rejmanek1979}} & Europe & terrestrial &       & No    & Host-parasitoid web &  \\
    146   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Rejmanek1979}} & Europe & terrestrial &       & No    & Host-parasitoid web &  \\
    147   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Rejmanek1979}} & Europe & terrestrial &       & No    & Host-parasitoid web &  \\
    148   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Rejmanek1979}} & Europe & terrestrial &       & No    & Host-parasitoid web &  \\
    149   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Force1974}} & USA   & terrestrial &       & No    & Host-parasitoid web &  \\
    150   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Hopkins1984}} & UK    & terrestrial &       & No    & Host-parasitoid web &  \\
    151   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Richards1926}} & UK    & terrestrial & 51,20,04 N & Yes   &       &  \\
    152   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Whittaker1984}} & USA   & terrestrial &       & No    & Source web &  \\
    153   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Mayse1978}} & USA   & terrestrial & 40,06,38 N & Yes   &       &  \\
    154   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Mayse1978}} & USA   & terrestrial & 39,56,41 N & Yes   &       &  \\
    155   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Askew1975}} & UK    & terrestrial &       & No    & Host-parasitoid web &  \\
    156   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Askew1961}} & UK    & terrestrial &       & No    & Host-parasitoid web &  \\
    157   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Askew1961}} & UK    & terrestrial &       & No    & Host-parasitoid web &  \\
    158   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Askew1961}} & UK    & terrestrial &       & No    & Host-parasitoid web &  \\
    159   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Askew1961}} & UK    & terrestrial &       & No    & Host-parasitoid web &  \\
    160   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Askew1961}} & UK    & terrestrial &       & No    & Host-parasitoid web &  \\
    161   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Askew1961}} & UK    & terrestrial &       & No    & Host-parasitoid web &  \\
    162   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Askew1961}} & UK    & terrestrial &       & No    & Host-parasitoid web &  \\
    163   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Askew1961}} & UK    & terrestrial &       & No    & Host-parasitoid web &  \\
    164   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Askew1961}} & UK    & terrestrial &       & No    & Host-parasitoid web &  \\
    165   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Askew1961}} & UK    & terrestrial &       & No    & Host-parasitoid web &  \\
    166   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Askew1961}} & UK    & terrestrial &       & No    & Host-parasitoid web &  \\
    167   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Askew1961}} & UK    & terrestrial &       & No    & Host-parasitoid web &  \\
    168   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Hawkins1984}} & USA   & terrestrial &       & No    & Host-parasitoid web &  \\
    169   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Hawkins1984}} & USA   & terrestrial &       & No    & Host-parasitoid web &  \\
    170   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Hawkins1984}} & USA   & terrestrial &       & No    & Host-parasitoid web &  \\
    171   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Hawkins1984}} & USA   & terrestrial &       & No    & Host-parasitoid web &  \\
    172   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Hawkins1984}} & USA   & terrestrial &       & No    & Host-parasitoid web &  \\
    173   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Hawkins1984}} & USA   & terrestrial &       & No    & Host-parasitoid web &  \\
    174   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Hawkins1984}} & USA   & terrestrial &       & No    & Host-parasitoid web &  \\
    175   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Hawkins1984}} & USA   & terrestrial &       & No    & Host-parasitoid web &  \\
    176   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Hawkins1984}} & USA   & terrestrial &       & No    & Host-parasitoid web &  \\
    177   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Hawkins1984}} & USA   & terrestrial &       & No    & Host-parasitoid web &  \\
    178   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Hawkins1984}} & USA   & terrestrial &       & No    & Host-parasitoid web &  \\
    179   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Robinson1953}} & England & terrestrial &       & No    & Source web &  \\
    180   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Savely1939}} & USA   & terrestrial & 36,01,02 N & Yes   &       &  \\
    181   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Savely1939}} & USA   & terrestrial & 36,01,02 N & Yes   &       &  \\
    182   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Savely1939}} & USA   & terrestrial & 36,01,02 N & Yes   &       &  \\
    183   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Savely1939}} & USA   & terrestrial & 36,01,02 N & Yes   &       &  \\
    184   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Beaver1972}} & UK    & terrestrial &       & No    & Could not locate original source &  \\
    185   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Chapman1955}} & UK    & terrestrial &       & No    & Source web &  \\
    186   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cornaby1974}} & Costa Rica & terrestrial &       & No    & Source web &  \\
    187   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cornaby1974}} & Costa Rica & terrestrial &       & No    & Source web &  \\
    188   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Jiron1981}} & Costa Rica & terrestrial &       & No    & Source web &  \\
    189   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{McKinnerney1978}} & USA   & terrestrial &       & No    & Could not locate original source &  \\
    190   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{McKinnerney1978}} & USA   & terrestrial &       & No    & Could not locate original source &  \\
    191   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{McKinnerney1978}} & USA   & terrestrial &       & No    & Could not locate original source &  \\
    192   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{McKinnerney1978}} & USA   & terrestrial &       & No    & Could not locate original source &  \\
    193   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{McKinnerney1978}} & USA   & terrestrial &       & No    & Could not locate original source &  \\
    194   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{McKinnerney1978}} & USA   & terrestrial &       & No    & Could not locate original source &  \\
    195   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Schoenly1983a}} & USA   & terrestrial &       & No    & Could not locate original source &  \\
    196   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Schoenly1983a}} & USA   & terrestrial &       & No    & Could not locate original source &  \\
    197   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Mohr1943}} & USA   & terrestrial &       & No    & Source web &  \\
    198   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Schoenly1983}} & Mexico & terrestrial &       & No    & Source web &  \\
    199   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Valiela1969}} & USA   & terrestrial &       & No    & Source web &  \\
    200   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Valiela1974}} & USA   & terrestrial &       & No    & Source web &  \\
    201   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Valiela1974}} & USA   & terrestrial &       & No    & Source web &  \\
    202   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Allan1982}} & USA   & stream & 38,50,21 N & Yes   &       &  \\
    203   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Collins1976}} & USA   & stream & 44,24 N & Yes   &       &  \\
    204   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Fryer1959}} & Malawi & lake  & 11,36,39 S & Yes   &       & Used whole-lake web only \\
    205   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Hildrew1985}} & UK    & stream & 51,01,05 N & Yes   &       &  \\
    206   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Jones1949}} & UK    & stream & 51,41 - 51,55 N & Yes   &       &  \\
    207   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Koslucher1973}} & USA   & stream & 41,57 - 42,14 N & Yes   &       &  \\
    208   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Minckley1963}} & USA   & stream & 37,56 N & Yes   &       &  \\
    209   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Minshall1967}} & USA   & stream & 37,18,26 N & Yes   &       &  \\
    210   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Percival1929}} & UK    & stream &       & No    & Generalized scheme &  \\
    211   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Ricker1935}} & Canada & stream &       & No    & Could not locate original source &  \\
    212   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Ricker1935}} & Canada & stream &       & No    & Could not locate original source &  \\
    213   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Badcock1949}} & UK    & stream & 53 N  & Yes   &       &  \\
    214   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2003}} & USA   & stream & 44,40,44 N & Yes   &       &  \\
    215   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2003}} & USA   & stream & 40,40,52 N & Yes   &       &  \\
    216   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2003}} & USA   & stream & 35,02,50 N & Yes   &       &  \\
    217   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2003}} & USA   & stream & 35,03,50 N & Yes   &       &  \\
    218   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 45.79 S & Yes   &       &  \\
    219   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 44.64 S & Yes   &       &  \\
    220   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 46.29 S & Yes   &       &  \\
    221   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 46.41 S & Yes   &       &  \\
    222   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 45,57,0 S & Yes   &       &  \\
    223   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 46.08 S & Yes   &       &  \\
    224   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 46.08 S & Yes   &       &  \\
    225   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 46.115 S & Yes   &       &  \\
    226   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 45.61 S & Yes   &       &  \\
    227   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 44.97 S & Yes   &       &  \\
    228   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 44.95 S & Yes   &       &  \\
    229   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 44.96 S & Yes   &       &  \\
    230   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 45 S  & Yes   &       &  \\
    231   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 45.58 S & Yes   &       &  \\
    232   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 45.61 S & Yes   &       &  \\
    233   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 45.782 S & Yes   &       &  \\
    234   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 45.803 S & Yes   &       &  \\
    235   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 45.752 S & Yes   &       &  \\
    236   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 46.29 S & Yes   &       &  \\
    237   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 46.41 S & Yes   &       &  \\
    238   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 45,57,0 S & Yes   &       &  \\
    239   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 46.08 S & Yes   &       &  \\
    240   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 46.08 S & Yes   &       &  \\
    241   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 46.115 S & Yes   &       &  \\
    242   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 46.29 S & Yes   &       &  \\
    243   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 46.41 S & Yes   &       &  \\
    244   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 45,57,0 S & Yes   &       &  \\
    245   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 46.08 S & Yes   &       &  \\
    246   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 46.08 S & Yes   &       &  \\
    247   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004c}} & New Zealand & stream & 46.115 S & Yes   &       &  \\
    248   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Thompson2004}} & New Zealand & estuary & 45,51,28 S & Yes   &       & Parasites removed \\
    249   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Angelini2005}} & Brazil & stream & 24,30 S & Yes   &       &  \\
    250   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Angelini2006}} & Brazil & lake  & 22,57,38 S & Yes   &       &  \\
    251   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Angelini2006}} & Brazil & lake  & 22,57,38 S & Yes   &       &  \\
    252   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Angelini2010}} & Brazil & stream & 18,40 S & Yes   &       &  \\
    253   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Angelini2011}} & Angola & marine & 5 - 17,15 S & Yes   &       &  \\
    254   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Angelini2013}} & Brazil & stream & 21 S  & Yes   &       &  \\
    255   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Angelini2013}} & Brazil & stream & 21 S  & Yes   &       &  \\
    256   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Angelini2013}} & Brazil & stream &       & No    & Could not locate original source &  \\
    257   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Baeta2011}} & Portugal & estuary & 40,08 N & Yes   &       &  \\
    258   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Baeta2011}} & Portugal & estuary & 40,08 N & Yes   &       &  \\
    259   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Baeta2011}} & Portugal & estuary & 40,08 N & Yes   &       &  \\
    260   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Baeta2011}} & Portugal & estuary & 40,08 N & Yes   &       &  \\
    261   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Baeta2011}} & Portugal & estuary & 40,08 N & Yes   &       &  \\
    262   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Baeta2011}} & Portugal & estuary & 40,08 N & Yes   &       &  \\
    263   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Schnieder1997}} & USA   & lake  & 45,58,50 N & Yes   &       &  \\
    264   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Stagliano2002}} & USA   & terrestrial & 39,06,27 N & Yes   &       &  \\
    265   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Lin2006}} & Taiwan & marine & 22,27 N & Yes   &       &  \\
    266   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cornejo-Donoso2008}} & Antarctica & marine & 60 - 69 S & Yes   &       &  \\
    267   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Zetina-Rejon2003}} & Mexico & marine & 23.15 - 22.9N & Yes   &       &  \\
    268   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cruz-Escalona2007}} & Mexico & marine & 18,45 - 18,50 N & Yes   &       &  \\
    269   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Liu2007}} & China & lake  & 29,32,35 N & Yes   &       &  \\
    270   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Filguera2011}} & Spain & marine & 42,17,37 - 42,19,29 N & Yes   &       &  \\
    271   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Amundsen2013}} & Norway & lake  & 69,07 N & Yes   &       & Parasites removed \\
    272   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Hampton2011}} & USA   & lake  & 48,03,29 N & Yes   &       &  \\
    273   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Parker2006}} & Alaska & stream & 69,1 N & Yes   &       &  \\
    274   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Parker2006}} & Alaska & stream & 69,1 N & Yes   &       &  \\
    275   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Parker2006}} & Alaska & stream & 69,1 N & Yes   &       &  \\
    276   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Parker2006}} & Alaska & stream & 69,1 N & Yes   &       &  \\
    277   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Massana1996}} & Spain & lake  & 42,08,6 - 42,07 N & Yes   &       &  \\
    278   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Stewart2011}} & USA   & lake  & 43.7 N & Yes   &       &  \\
    279   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Stewart2011}} & USA   & lake  & 43.7 N & Yes   &       &  \\
    280   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cromar1996}} & Canada & stream & 43,58 N & Yes   &       &  \\
    281   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cromar1996}} & Canada & stream & 44,58 N & Yes   &       &  \\
    282   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cromar1996}} & Canada & stream & 45,58 N & Yes   &       &  \\
    283   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cromar1996}} & Canada & stream & 46,58 N & Yes   &       &  \\
    284   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cromar1996}} & Canada & stream & 47,58 N & Yes   &       &  \\
    285   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cromar1996}} & Canada & stream & 48,58 N & Yes   &       &  \\
    286   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cromar1996}} & Canada & stream & 49,58 N & Yes   &       &  \\
    287   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cromar1996}} & Canada & stream & 50,58 N & Yes   &       &  \\
    288   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Christian1999}} & USA   & estuary & 30,05,39 N & Yes   &       &  \\
    289   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Fetahi2011}} & Ethiopia & lake  & 11,15 N & Yes   &       &  \\
    290   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Brodeur1992}} & USA   & marine & 43 - 48 N & Yes   &       &  \\
    291   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Brodeur1992}} & USA   & marine & 44 - 48 N & Yes   &       &  \\
    292   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Brodeur1992}} & USA   & marine & 45 - 48 N & Yes   &       &  \\
    293   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Brodeur1992}} & USA   & marine & 46 - 48 N & Yes   &       &  \\
    294   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Gaedke1994} & Europe & lake  & 47,37,30 N & Yes   &       &  \\
    295   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Preston2012}} & USA   & lake  & 37.66 N & Yes   &       & Parasites removed \\
    296   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Ratsirarson1996}} & Madagascar & lake  & 24,55 S & Yes   &       &  \\
    297   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Closs1994}} & Australia & stream & 37,40 S & Yes   &       &  \\
    298   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Closs1994}} & Australia & stream & 37,40 S & Yes   &       &  \\
    299   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Closs1994}} & Australia & stream & 37,40 S & Yes   &       &  \\
    300   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Closs1994}} & Australia & stream & 37,40 S & Yes   &       &  \\
    301   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Closs1994}} & Australia & stream & 37,40 S & Yes   &       &  \\
    302   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Closs1994}} & Australia & stream & 37,40 S & Yes   &       &  \\
    303   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Closs1994}} & Australia & stream & 37,40 S & Yes   &       &  \\
    304   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Closs1994}} & Australia & stream & 37,40 S & Yes   &       &  \\
    305   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Closs1994}} & Australia & stream & 37,40 S & Yes   &       &  \\
    306   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Closs1994}} & Australia & stream & 37,40 S & Yes   &       &  \\
    307   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Closs1994}} & Australia & stream & 37,40 S & Yes   &       &  \\
    308   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Closs1994}} & Australia & stream & 37,40 S & Yes   &       &  \\
    309   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Gontikaki2011}} & Atlantic Ocean & marine & 61,13,88 N & Yes   &       &  \\
    310   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Khan2009}} & India & lake  & 12,45,25 N & Yes   &       &  \\
    311   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Memmott2000}} & UK    & terrestrial &       & No    & Source web &  \\
    312   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Kaiser-Bunbury2009}} & Mauritius & terrestrial &       & No    & Plant-pollinator web &  \\
    313   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Kaiser-Bunbury2009}} & Mauritius & terrestrial &       & No    & Plant-pollinator web &  \\
    314   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Kaiser-Bunbury2011}} & Mauritius & terrestrial &       & No    & Plant-pollinator web &  \\
    315   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Kaiser-Bunbury2011}} & Mauritius & terrestrial &       & No    & Plant-pollinator web &  \\
    316   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Kaiser-Bunbury2011}} & Mauritius & terrestrial &       & No    & Plant-pollinator web &  \\
    317   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Kaiser-Bunbury2011}} & Mauritius & terrestrial &       & No    & Plant-pollinator web &  \\
    318   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Kaiser-Bunbury2011}} & Mauritius & terrestrial &       & No    & Plant-pollinator web &  \\
    319   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Kaiser-Bunbury2011}} & Mauritius & terrestrial &       & No    & Plant-pollinator web &  \\
    320   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Ruzicka2012}} & USA   & marine & 42 - 49 N & Yes   &       &  \\
    321   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Ruzicka2012}} & USA   & marine & 43 - 49 N & Yes   &       &  \\
    322   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Ruzicka2012}} & USA   & marine & 44 - 49 N & Yes   &       &  \\
    323   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Ruzicka2012}} & USA   & marine & 45 - 49 N & Yes   &       &  \\
    324   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Ruzicka2012}} & USA   & marine & 46 - 49 N & Yes   &       &  \\
    325   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Lewis2002}} & Belize & terrestrial &       & No    & Host-parasitoid web &  \\
    326   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Lewis2002}} & Belize & terrestrial &       & No    & Host-parasitoid web &  \\
    327   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Kelleway2010}} & Australia & stream & 29,13 S & Yes   &       &  \\
    328   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Kelleway2010}} & Australia & stream & 29,20,0 S & Yes   &       &  \\
    329   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Kelleway2010}} & Australia & stream & 29,35 S & Yes   &       &  \\
    330   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Kelleway2010}} & Australia & stream & 29,13 S & Yes   &       &  \\
    331   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Kelleway2010}} & Australia & stream & 29,20,0 S & Yes   &       &  \\
    332   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Huang2008}} & China & stream & 31.70 N & Yes   &       &  \\
    333   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Link2002}} & USA   & marine & 33,51 - 44,49 N & Yes   &       &  \\
    334   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Alcorlo2001}} & Spain & lake  & 41,26,36 N & Yes   &       &  \\
    335   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Alcorlo2001}} & Spain & lake  & 41,26,05 N & Yes   &       &  \\
    336   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Alcorlo2001}} & Spain & lake  & 41,26,36 N & Yes   &       &  \\
    337   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Alcorlo2001}} & Spain & lake  & 41,26,05 N & Yes   &       &  \\
    338   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Torres2013}} & South Atlantic & marine &       & No    & Could not locate original source &  \\
    339   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Bulman2001}} & Australia & marine & 37-39 S & Yes   &       &  \\
    340   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Smith2007}} & Antarctica & marine & 74-78 S & Yes   &       &  \\
    341   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Lancraft1991}} & Antarctica & marine & 58,20.8 - 60,10.4 S & Yes   &       &  \\
    342   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Hopkins1993}} & Antarctica & marine & 57,00 - 60,45 S & Yes   &       &  \\
    343   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Kitching1987}} & Australia & lake  & 28,12 S & Yes   &       &  \\
    344   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Hodkinson2004}} & Norway & terrestrial &       & No    & Host-parasitoid web &  \\
    345   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Boit2012}} & Europe & lake  & 47,38,10 N & Yes   &       &  \\
    346   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Rayner2010}} & Australia & stream & 17,7 S & Yes   &       &  \\
    347   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Motta2005}} & Brazil & stream & 23,03 S & Yes   &       &  \\
    348   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Douglass2011}} & USA   & marine & 37,13 N & Yes   &       &  \\
    349   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Douglass2011}} & USA   & marine & 37,13 N & Yes   &       &  \\
    350   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Douglass2011}} & USA   & marine & 37,13 N & Yes   &       &  \\
    351   & ~\textbackslash{}citet{GlobalWeb} & ~\textbackslash{}citet{Warren1989} & UK    & lake  &       & No    & Could not locate original source &  \\
    352   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Poepperl2003}} & Germany & stream & 54,06 N & No    & describes compartments, not species &  \\
    353   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Poepperl2003}} & Germany & stream & 54,06 N & No    & describes compartments, not species &  \\
    354   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Goldwasser1993}} & St. Martin & terrestrial & 18,03 N & Yes   &       &  \\
    355   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cohen2003}} & USA   & lake  & 46,13 N & Yes   &       &  \\
    356   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Cohen2003}} & USA   & lake  & 46,13 N & Yes   &       &  \\
    357   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Yodzis1998}} & South Africa & marine & 34,21 - 16 S & No    & Study site unclear &  \\
    358   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Schroter2003}} & Europe & terrestrial &       & No    & Generalized scheme &  \\
    359   & ~\textbackslash{}citet{GlobalWeb} & \multicolumn{1}{l}{~\textbackslash{}citet{Baiser2012}} & North America & lake  &       & No    & Study site unclear &  \\
    \bottomrule
    \end{tabular}%
  \label{tab:addlabel}%
\end{table}%