### 2015-09-07 manuscript/supplement.tex
Maybe done?

\documentclass[12pt]{article}  
\usepackage{amsmath}
\usepackage{gensymb}
\usepackage{url}
\usepackage[dvips]{graphicx}
\usepackage{multirow}
\usepackage{geometry}
\usepackage{pdflscape}
\usepackage[labelfont=bf]{caption}
\usepackage{setspace}
\usepackage[numbers,sort&compress]{natbib}
\usepackage{float}
\usepackage[running]{lineno}

% some cheats to reduce the need to type complicated bits and pieces
\newcommand{\expect}[1]{\left\langle #1 \right\rangle}
\newcommand{\etal}{\textit{et al.\ }}

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

% the abstract formatting
\newenvironment{sciabstract}{%
\begin{quote} \bf}
{\end{quote}}

\renewcommand\refname{References}
% the abstract formatting

% margin sizes`
\topmargin 0.0cm
\oddsidemargin 0.2cm

% let's get some nature formatted citations
%\usepackage{overcite}
\textwidth 16cm 
\textheight 21cm
\footskip 1.0cm

\begin{document}

\title{Supplementary Information}
\author{Alyssa R. Cirtwill$^{1,2}$, Daniel B. Stouffer$^{1}$, Tamara N. Romanuk$^{2}$}
\date{\footnotesize$^1$Centre for Integrative Ecology\\School of Biological Sciences\\University of Canterbury\\
Private Bag 4800\\Christchurch 8140, New Zealand \\
\medskip$^2$Department of Biology\\
Life Science Centre, Dalhousie University\\1355 Oxford St., P0 BOX 15000\\
Halifax NS, B3H 4R2, Canada\\}

\maketitle
\baselineskip=8.5mm
% \linenumbers

\vspace{0.4 in}
\beginsupplement
\linenumbers
% \begin{landscape}

\subsection*{Appendix S1: details of food web sources and selection}

  We combined published food webs from~\cite{GlobalWeb},~\cite{Riede2011},
  and~\cite{Dunne2013}. Of the 358 food webs available
  from~\cite{GlobalWeb}, 59 were rejected because their original source was
  unpublished or could not be retrieved. We further eliminated 13 source or
  sink webs (which describe feeding links flowing from or to a particular
  resource or top predator rather than the interactions of an entire
  community), 8 plant-pollinator webs, 33 host-parasitoid webs, and 2 networks
  describing competitive interactions. Finally, we eliminated 2 webs
  describing  inferred interactions in an extinct community, 10 ``generalised
  schemes'' that were not based on direct field data, and 4 webs where it was
  not clear which of several described study sites were used to construct the
  food web (and therefore the latitude of the food web could not be
  included). This left us with 226 webs from~\cite{GlobalWeb}.

  To these, we added 7 webs from~\cite{Dunne2013} that were originally used
  to assess the roles of parasites within food webs. As all but 3 of the webs
  from~\cite{GlobalWeb} contained only free-living species, we removed the
  the parasites (and all interactions involving them) from these webs (and the
  3 webs containing parasites from~\cite{GlobalWeb}) leaving free-living
  species only. In addition, we included 30 of the 40 webs
  from~\cite{Riede2011}. Nine of the
  remaining webs in the~\cite{Riede2011} dataset were also included in
  the~\cite{GlobalWeb}  dataset, and the last web was derived from an
  unpublished source. This gave us a database of 263 webs. As a measure of
  quality-control, we then eliminated any food web which did not have at least
  one each of basal resources, intermediate consumers, and top predators. Webs
  which did not meet this requirement tended to be very small and poorly-resolved.
  Our final sample size was thereby reduced to 196 food webs.

  % Tree hole and pitcher plant webs counted as "lakes", wetlands could be either lakesor streams depending on the study

 See Table~\ref{S1} for a list of rejected webs and reasons for their
  exclusion.

  \newpage
  \begin{landscape}

    \begin{table}[h!]
    \centering
    \caption{Sources of food webs considered for use in this study. ``Collection" indicates the direct source of the
    food web (i.e., where the interaction matrix was obtained) while ``Original source" indicates the original publication
    of the food web. We also show the ecosystem type, country of origin, and latitude for each food web, as well as whether
    it was ultimately included in the study. Collection 1 is~\cite{Riede2011}, collection 2 is~\cite{Dunne2013}, and collection 3 is~\cite{GlobalWeb}.}
    \label{S1}

    {\footnotesize
      \begin{tabular}{p{2.8cm}p{1.3cm}p{5.5cm}p{2.2cm}p{2.5cm}lp{3.5cm}}
        \hline
        Web ID & Collection & Original Source & Ecosystem & Latitude(degrees) & Used  & Reason for rejection \\
        \hline
        Alamitos & 1 & \cite{Harrison1995}    & stream & 37,12,25 N & Yes   &       \\
        Alford & 1 & \cite{Havens1992}    & lake  & 44,15,41.59 N & Yes   &       \\
        Beaver & 1 & \cite{Havens1992}    & lake  & 44,18,58.13 N & Yes   &       \\
        Bere  & 1 & \cite{Woodward2008}  & stream & 50,45,7 N & Yes   &       \\
        Bridge & 1 & \cite{Havens1992}    & lake  & 44,09,42.66 N & Yes   &       \\
        Broad & 1 & \cite{Townsend1998}  & stream & 45.896938 S & Yes   &       \\
        Broadstone & 1 & \cite{Woodward2005} & stream  & 51,01,05 N & Yes   &       \\
        Calero & 1 & \cite{Harrison1995}    & stream & 37,11,09 - 20 N & Yes   &       \\
        Canton & 1 & \cite{Townsend1998}  & stream & 45.752 S & No    & Duplicated in GlobalWeb (web 235) \\
        Carpinteria2006 & 1 & \cite{Lafferty2006}    & estuary & 34.41 - 34.40 N & Yes   &       \\
        Chub  & 1 & \cite{Havens1992}    & lake  & 43,31,39.36 N & Yes   &       \\
        Coachella & 1 & \cite{Polis1991a} & terrestrial & 33,54 N & Yes   &       \\
        Connery & 1 & \cite{Havens1992}    & lake  & 44,18,42.10 N & Yes   &       \\
        Corte & 1 & \cite{Harrison1995}    & stream & 37,56,28 - 37,57,39 N & Yes   &       \\
        Coyote & 1 & \cite{Harrison1995}    & stream & 37,53,11.66 - 37,53,17.06 N & Yes   &       \\
        German & 1 & \cite{Townsend1998}  & stream & 44.97 S & No    & Duplicated in GlobalWeb (web 227) \\
        GrandeCl & 1 & \cite{CattinBlandenier2004} & terrestrial & 46,47,09 N & Yes   &       \\
        GrandeSc & 1 & \cite{CattinBlandenier2004} & terrestrial & 46,47,09 N & Yes   &       \\
        Guadalupe & 1 & \cite{Harrison1995}    & stream & 37,27,54 - 37,21,17 N & Yes   &       \\
        Hainich   & 1 & Unpublished  & terrestrial &       & No    & Could not locate original source \\
        LittleKyeBurn & 1 & \cite{Townsend1998}  & stream & 45 S  & No    & Duplicated in GlobalWeb (web 230) \\
        LittleRock & 1 & \cite{Martinez1991}    & lake  & 45,44,07.94 N & Yes   &       \\
        Mondego & 1 & \cite{Patricio2006}  & estuary & 40,08 N & Yes   &       \\
        \hline
      \end{tabular}}%
      \end{table}

        \newpage

    \begin{table}[h!]
    \centering
    {\footnotesize
      \begin{tabular}{p{2.8cm}p{1.3cm}p{5.5cm}p{2.2cm}p{2.5cm}lp{3.5cm}}
        \hline
        Web ID & Collection & Original Source & Ecosystem & Latitude(degrees) & Used  & Reason for rejection  \\
        \hline
        Montane Forest & 1 & \cite{Rasmussen1941}    & terrestrial & 36,5-37 N & No    & Duplicated in GlobalWeb (web60) \\
        Reef  & 1 & \cite{Opitz1996}  & marine & 18,29 - 17,53 N & Yes   &       \\
        Sierra & 1 & \cite{Brose2005}    & lake  & 36,50 - 37,02 N & Yes   &       \\
        SimberloffE1 & 1 & \cite{Simberloff1976}  & terrestrial & 24,40,37.29 N & Yes   &       \\
        SimberloffE2 & 1 & \cite{Simberloff1976}  & terrestrial & 24,40,33 N & Yes   &       \\
        SimberloffE3 & 1 & \cite{Simberloff1976}  & terrestrial & 24,41,23.82 N & Yes   &       \\
        SimberloffE7 & 1 & \cite{Simberloff1976}  & terrestrial & 25,15 N & Yes   &       \\
        SimberloffE9 & 1 & \cite{Simberloff1976}  & terrestrial & 25,01,59.64 - 25,02,55.44 N & Yes   &       \\
        SimberloffST2 & 1 & \cite{Simberloff1976}  & terrestrial & 24,36,16 N & Yes   &       \\
        Skipwith & 1 & \cite{Warren1989}  & lake  & 53,40 N & Yes   &       \\
        Stink & 1 & \cite{Havens1992}    & lake  & 43,37,55.88 N & Yes   &       \\
        StMarks & 1 & \cite{Christian1999}    & estuary & 30,05 N & No    & Duplicated in GlobalWeb (web 288) \\
        Stony  & 1 & \cite{Townsend1998}  & stream & 45.58 S & No    & Duplicated in GlobalWeb (web 231) \\
        TreleaseWoods & 1 & \cite{Twomey1945}    & terrestrial & 40,09,47 N & No    & Duplicated in GlobalWeb (web 59) \\
        Tuesday & 1 & \cite{Jonsson2005}    & lake  & 46,13 N & Yes   &       \\
        Weddell Sea & 1 & \cite{Brose2005}  & marine & 58,20.8 - 60,10.4 S & No    & Duplicated in GlobalWeb (web 341) \\
        Ythan & 1 & \cite{Huxham1996}     & marine &       & No    & Duplicated in Dunne (YthanDunne) \\
        BahiaDunne & 2 & \cite{Hechinger2011}  & estuary & 30.43 - 30.46 N & Yes   &        \\
        CarpinteriaDunne & 2 & \cite{Hechinger2011}    & estuary & 34.41 - 34.40 N & Yes   &        \\
        EsteroDunne & 2 & \cite{Hechinger2011}  & estuary & 31.78 - 31.69 N & Yes   &        \\
        FjordDunne & 2 & \cite{Zander2011}  & estuary & 54.92 - 54.75 N & Yes   &        \\
        OtagoDunne & 2 & \cite{Mouritsen2011}  & estuary & 45.78 - 45.88 S & Yes   &        \\
        SyltDunne & 2 & \cite{Thieltges2011}  & estuary & 54.83 - 55.12 N & Yes   &        \\
        YthanDunne & 2 & \cite{Huxham1996}  & estuary & 57,18,32 N & Yes   &        \\
        1     & 3 & \cite{Qazim1970}  & estuary &       & No    & Could not locate original source \\
        2     & 3 & \cite{Day1967} & estuary &       & No    & Could not locate original source \\
        3     & 3 & \cite{Woodwell1967}    & estuary & 41,03,11 N & Yes   &       \\
        \hline
      \end{tabular}}%
      \end{table}

        \newpage

    \begin{table}[h!]
    \centering
    {\footnotesize
      \begin{tabular}{p{2.8cm}p{1.3cm}p{5.5cm}p{2.2cm}p{2.5cm}lp{3.5cm}}
        \hline
        Web ID & Collection & Original Source & Ecosystem & Latitude(degrees) & Used  & Reason for rejection  \\
        \hline
        4     & 3 & \cite{Johnston1956}    & estuary &       & No    & Sink web \\
        5     & 3 & \cite{Teal1962}    & estuary & 31,28,38 N & Yes   &       \\
        6     & 3 & \cite{MacGinitie1935}    & estuary & 36,48,18 N & Yes   &       \\
        7     & 3 & \cite{Kremer1978}    & marine & 41,45 - 41,30 N & Yes   &       \\
        8     & 3 & \cite{Nixon1973}    & estuary & 41,33,00 N & Yes   &       \\
        9     & 3 & \cite{Kitching1967}  & stream &       & No    & Could not locate original source \\
        10    & 3 & \cite{Menge1976}    & marine & 41 - 47 N & Yes   &       \\
        11    & 3 & \cite{Menge1976}    & marine & 41 - 47 N & Yes   &       \\
        12    & 3 & \cite{Menge1976}    & marine & 45,45 - 49 N & Yes   &       \\
        13    & 3 & \cite{Menge1976}    & marine & 45,45 - 49 N & Yes   &       \\
        14    & 3 & \cite{Walsh1967}  & estuary & 21,26,09 N & No    & Could not locate original source \\
        15    & 3 & \cite{Walsh1967}  & estuary & 21,26,09 N & No    & Could not locate original source \\
        16    & 3 & \cite{Copeland1974}    & estuary &       & No    & Could not locate original source \\
        17    & 3 & \cite{Hiatt1960}  & marine & 4,30 - 15 N & No    & Generalised scheme \\
        18    & 3 & \cite{Niering1963a}  & terrestrial & 1,01,29 - 1,06,23 N & Yes   &       \\
        19    & 3 & \cite{Brooks1963}    & lake &       & No    & Could not locate original source \\
        20    & 3 & \cite{Knox1970}  & terrestrial &       & No    & Could not locate original source \\
        21    & 3 & \cite{Patten1979}  & marine & 70-85 S & Yes   &       \\
        22    & 3 & \cite{Summerhayes1923}  & terrestrial & 74,27,54 N & Yes   &       \\
        23    & 3 & \cite{Bird1930}  & terrestrial & 50,25,17 N & Yes   &       \\
        24    & 3 & \cite{Bird1930}  & terrestrial & 50,25,17 N & Yes   &       \\
        25    & 3 & \cite{Bird1930}  & terrestrial & 50,25,17 N & Yes   &       \\
        26    & 3 & \cite{Bird1930}  & terrestrial & 50,25,17 N & Yes   &       \\
        27    & 3 & \cite{Varley1970}  & terrestrial &       & No    & Could not locate original source \\
        28    & 3 & \cite{Paviour-Smith1956}  & estuary & 45,51,30 S & Yes   &       \\
        29    & 3 & \cite{Dunbar1953}  & marine &       & No    & Generalised scheme \\
        30    & 3 & \cite{Mackintosh1964}  & marine &       & No    & Generalised scheme \\
        31    & 3 & \cite{Petipa1970}  & marine &       & No    & Could not locate original source \\
      \hline
      \end{tabular}}%
      \end{table}

        \newpage

    \begin{table}[h!]
    \centering
    {\footnotesize
      \begin{tabular}{p{2.8cm}p{1.3cm}p{5.5cm}p{2.2cm}p{2.5cm}lp{3.5cm}}
        \hline
        Web ID & Collection & Original Source & Ecosystem & Latitude(degrees) & Used  & Reason for rejection  \\
        \hline
        32    & 3 & \cite{Petipa1970}  & marine &       & No    & Could not locate original source \\
        33    & 3 & \cite{Fryer1959} & lake & 11,36,39 S & Yes   &       \\
        34    & 3 & \cite{Jones1949}     & stream & 51,41 - 51,55 N & Yes   &       \\
        35    & 3 & \cite{Minshall1967}    & stream & 37,18,26 N & Yes   &       \\
        36    & 3 & \cite{Walsh1967}  & estuary & 21,26,09 N & No    & Could not locate original source \\
        37    & 3 & \cite{Clarke1967}    & marine & 41,02,49 N & Yes   &       \\
        38    & 3 & \cite{Fryer1959} & lake & 11,36,39 S & Yes   &       \\
        39    & 3 & \cite{Fryer1959} & lake & 11,36,39 S & Yes   &       \\
        40    & 3 & \cite{Harrison1962}  & terrestrial &       & No    & Generalised scheme \\
        41    & 3 & \cite{Parin1970}  & marine &       & No    & Could not locate original source \\
        42    & 3 & \cite{Vinogradov1978}  & marine &       & No    & Study site unclear \\
        43    & 3 & \cite{Rosenthal1974}    & marine & 32,57 N & Yes   &       \\
        44    & 3 & \cite{Yanez1978}  & marine & 17,26,21 N & Yes   &       \\
        45    & 3 & \cite{Tilly1968}    & stream & 41,22,55 N & Yes   &       \\
        46    & 3 & \cite{Patten1975}    & lake & 33,54 N & Yes   &       \\
        47    & 3 & \cite{Harris1980}    & lake &       & No    & Study site unclear \\
        48    & 3 & \cite{Simenstad1978}    & marine &       & No    & Extinct species \\
        49    & 3 & \cite{Simenstad1978}    & marine &       & No    & Extinct species \\
        50    & 3 & \cite{Nybakken1982}    & marine &       & No    & Generalised scheme \\
        51    & 3 & \cite{Dexter1947}    & marine & 42,38,56 N & Yes   &       \\
        52    & 3 & \cite{Paine1980a}    & marine & 58,27 N & Yes   &       \\
        53    & 3 & \cite{Paine1980a}    & marine & 48,19 - 48,37 N & Yes   &       \\
        54    & 3 & Unpublished  & marine &       & No    & Unpublished \\
        55    & 3 & \cite{Milne1972}  & estuary &       & No    & Could not locate original source \\
        56    & 3 & \cite{Milne1972}  & estuary &       & No    & Could not locate original source \\
        57    & 3 & \cite{Yanez1978}  & marine & 17,26,21 N & Yes   &       \\
        58    & 3 & \cite{Smirnov1961}  & lake & 51,46 N & Yes   &       \\
        59    & 3 & \cite{Twomey1945}    & terrestrial & 40,09,47 N & Yes   &       \\
        60    & 3 & \cite{Rasmussen1941}    & terrestrial & 36,5-37 N & Yes   &       \\
        61    & 3 & \cite{Summerhayes1928}  & terrestrial & 77,52,30 N & Yes   &       \\
        62    & 3 & \cite{Summerhayes1928}  & terrestrial & 77,52,30 N & Yes   &       \\
        \hline
      \end{tabular}}%
      \end{table}

        \newpage

    \begin{table}[h!]
    \centering
    {\footnotesize
      \begin{tabular}{p{2.8cm}p{1.3cm}p{5.5cm}p{2.2cm}p{2.5cm}lp{3.5cm}}
        \hline
        Web ID & Collection & Original Source & Ecosystem & Latitude(degrees) & Used  & Reason for rejection  \\
        \hline
        63    & 3 & \cite{Jones1950}     & stream & 52,25,08 N & Yes   &       \\
        64    & 3 & \cite{Cummins1966}    & stream & 41,39,33 N & Yes   &       \\
        65    & 3 & \cite{Tsuda1972}  & stream & 34,21,20 - 34,23,21 N & Yes   &       \\
        66    & 3 & \cite{Mann1972a}  & stream & 51,30 N & Yes   &       \\
        67    & 3 & \cite{Carlson1968}    & stream & 40,24,16 N & Yes   &       \\
        68    & 3 & \cite{Morgan1972}  & lake &       & No    & Could not locate original source \\
        69    & 3 & \cite{Cohen1990}  & estuary &       & No    & Could not locate original source \\
        70    & 3 & \cite{Kemp1977}    & estuary & 28,54,08 N & Yes   &       \\
        71    & 3 & \cite{Sorokin1972}  & lake & 48,00 N & Yes   &       \\
        72    & 3 & \cite{Baril1983}  & lake & 46,02 N & Yes   &       \\
        73    & 3 & \cite{Schiemer1979}  & lake &       & No    & Could not locate original source \\
        74    & 3 & \cite{Brown1964} & marine & 34,15,00 S & Yes   &       \\
        75    & 3 & \cite{Pechlaner1972}  & lake & 47,11 N & Yes   &       \\
        76    & 3 & \cite{Schiemer1979}  & lake & 47,42 - 47,54 N & Yes   &       \\
        77    & 3 & \cite{Cohen1990}  & lake &       & No    & Could not locate original source \\
        78    & 3 & \cite{Burgis1972}  & lake & 0     & Yes   &       \\
        79    & 3 & \cite{Sarvala1974}  & lake &       & No    & Could not locate original source \\
        80    & 3 & \cite{Sarvala1974}  & lake &       & No    & Could not locate original source \\
        81    & 3 & \cite{Hatanaka1977}  & marine &       & No    & Could not locate original source \\
        82    & 3 & \cite{Cohen1990} & marine &       & No    & Could not locate original source \\
        83    & 3 & \cite{Cohen1990}  & lake &       & No    & Could not locate original source \\
        84    & 3 & \cite{Wilbur1972}    & lake & 42,28,42 N & Yes   &       \\
        85    & 3 & \cite{Mizuno1982}  & lake &       & No    & Could not locate original source \\
        86    & 3 & \cite{Hogetsu1979}  & marine & 34,36 - 35,38 N & Yes   &       \\
        87    & 3 & \cite{Bradstreet1982}  & terrestrial & 72,46,34.95 N & Yes   &       \\
        \hline
      \end{tabular}}%
      \end{table}

        \newpage

    \begin{table}[h!]
    \centering
    {\footnotesize
      \begin{tabular}{p{2.8cm}p{1.3cm}p{5.5cm}p{2.2cm}p{2.5cm}lp{3.5cm}}
        \hline
        Web ID & Collection & Original Source & Ecosystem & Latitude(degrees) & Used  & Reason for rejection  \\
        \hline
        88    & 3 & \cite{Kuusela1980}  & stream &       & No    & Could not locate original source \\
        89    & 3 & \cite{Hartley1948}  & stream & 52,13 N & Yes   &       \\
        91    & 3 & \cite{Cohen1990}  & terrestrial &       & No    & Could not locate original source \\
        92    & 3 & \cite{Swan1961}  & terrestrial &       & No    & Generalised scheme \\
        93    & 3 & \cite{Pattie1966}    & terrestrial & 45 N  & Yes   &       \\
        94    & 3 & \cite{Brown1971}    & terrestrial &       & No    & Could not locate original source \\
        95    & 3 & \cite{Brown1975}    & terrestrial & 70,11,14.18 N & Yes   &       \\
        96    & 3 & \cite{Cohen1990}  & terrestrial &       & No    & Could not locate original source \\
        97    & 3 & \cite{Cohen1990}  & terrestrial &       & No    & Could not locate original source \\
        98    & 3 & \cite{Holm1980}  & terrestrial &       & No    & Could not locate original source \\
        99    & 3 & \cite{Howes1954}    & terrestrial & 32,36 N & Yes   &       \\
        100   & 3 & \cite{Cohen1990}  & terrestrial &       & No    & Could not locate original source \\
        101   & 3 & \cite{Cohen1990} & lake  &       & No    & Could not locate original source \\
        102   & 3 & \cite{Shushkina1979}  & marine & 10 S - 15 N & Yes   &       \\
        103   & 3 & \cite{Petipa1979}  & marine &       & No    & Generalised scheme \\
        104   & 3 & \cite{Menge1976}  & marine & 4,48,04 N & Yes   &       \\
        105   & 3 & \cite{Edwards1982}  & marine & 42,40 N & Yes   &       \\
        106   & 3 & \cite{Glynn1965}  & marine &       & No    & Could not locate original source \\
        107   & 3 & \cite{Peterson1979}  & marine & 39,52,02 N & Yes   &       \\
        108   & 3 & \cite{Hewatt1937}  & marine & 36,37 N & Yes   &       \\
        109   & 3 & \cite{Castilla1981} & marine &       & No    & Could not locate original source \\
        110   & 3 & \cite{Dexter1947}  & marine & 42,38,56 N & Yes   &       \\
        111   & 3 & \cite{Dexter1947}  & marine & 42,38,56 N & Yes   &       \\
        112   & 3 & \cite{Dexter1947}  & marine & 42,38,56 N & Yes   &       \\
        113   & 3 & \cite{Dexter1947}  & marine & 42,38,56 N & Yes   &       \\
        114   & 3 & \cite{Marshall1982}  & lake  & 17,54,11.90 S & Yes   &       \\
        \hline
      \end{tabular}}%
      \end{table}

        \newpage

    \begin{table}[h!]
    \centering
    {\footnotesize
      \begin{tabular}{p{2.8cm}p{1.3cm}p{5.5cm}p{2.2cm}p{2.5cm}lp{3.5cm}}
        \hline
        Web ID & Collection & Original Source & Ecosystem & Latitude(degrees) & Used  & Reason for rejection  \\
        \hline
        115   & 3 & \cite{Koepcke1952}  & marine &       & No    & Could not locate original source \\
        116   & 3 & \cite{Hurlbert1972}  & lake  & 35,09,19 S & Yes   &       \\
        117   & 3 & \cite{Zaret1973}  & lake  & 9,11,31.50 N & Yes   &       \\
        118   & 3 & \cite{Bindloss1972}  & lake  & 56,41,18.71 N & Yes   &       \\
        119   & 3 & \cite{Mann1972}  & stream & 51,27 N & Yes   &       \\
        120   & 3 & \cite{Moriarty1973}  & lake  & 0,00,58 N & Yes   &       \\
        121   & 3 & \cite{VanEs1977}  & estuary & 53,20 - 53,15 N & Yes   &       \\
        122   & 3 & \cite{Nixon1973}  & estuary & 41,33,04 N & Yes   &       \\
        123   & 3 & \cite{Harris1972}  & terrestrial & 40,45,28 N & Yes   &       \\
        124   & 3 & \cite{Collins1976}  & stream & 44,34 N & Yes   &       \\
        125   & 3 & \cite{Landry1977}  & marine &       & No    & Generalised scheme \\
        126   & 3 & \cite{Larsson1978}  & lake  & 61,26,00 N & Yes   &       \\
        127   & 3 & \cite{Beaver1979}  & lake  &       & No    & Could not locate original source \\
        128   & 3 & \cite{Beaver1979}  & lake  &       & No    & Could not locate original source \\
        129   & 3 & Unpublished & terrestrial &       & No    & Unpublished \\
        130   & 3 & \cite{Odum1975}  & estuary & 27,12 N & Yes   &       \\
        131   & 3 & \cite{Beaver1985}  & lake  & 4,12,37 N & Yes   &       \\
        132   & 3 & \cite{Beaver1985}  & lake  & 4,12,37 N & Yes   &       \\
        133   & 3 & \cite{Beaver1985}  & lake  & 7,52,22 N & Yes   &       \\
        134   & 3 & \cite{Beaver1985}  & lake  & 19 S  & Yes   &       \\
        135   & 3 & \cite{Beaver1985}  & lake  & 4,40,46 S & Yes   &       \\
        136   & 3 & \cite{Bradshaw1983}    & lake  &       & No    & Could not locate original source \\
        137   & 3 & \cite{Corker1984}  & lake  & 22,14-22,17 N & Yes   &       \\
        138   & 3 & \cite{Kitching1983}     & lake  &       & No    & Could not locate original source \\
        139   & 3 & \cite{Kitching1983}  & lake  &       & No    & Could not locate original source \\
        140   & 3 & \cite{Seifert1979}  & terrestrial & 10,21 N & Yes   &       \\
        141   & 3 & \cite{Seifert1976}  & terrestrial &       & No    & Competition network \\
        142   & 3 & \cite{Seifert1976}  & terrestrial &       & No    & Competition network \\
        143   & 3 & \cite{Snow1958}    & lake  & 34,41,45 N & Yes   &       \\
         \hline
      \end{tabular}}%
      \end{table}

        \newpage

    \begin{table}[h!]
    \centering
    {\footnotesize
      \begin{tabular}{p{2.8cm}p{1.3cm}p{5.5cm}p{2.2cm}p{2.5cm}lp{3.5cm}}
        \hline
        Web ID & Collection & Original Source & Ecosystem & Latitude(degrees) & Used  & Reason for rejection  \\
        \hline
        144   & 3 & \cite{Snow1958}  & lake  &       & No    & Could not locate original source \\
        145   & 3 & \cite{Rejmanek1979}  & terrestrial &       & No    & Host-parasitoid web \\
        146   & 3 & \cite{Rejmanek1979}  & terrestrial &       & No    & Host-parasitoid web \\
        147   & 3 & \cite{Rejmanek1979}  & terrestrial &       & No    & Host-parasitoid web \\
        148   & 3 & \cite{Rejmanek1979}  & terrestrial &       & No    & Host-parasitoid web \\
        149   & 3 & \cite{Force1974}    & terrestrial &       & No    & Host-parasitoid web \\
        150   & 3 & \cite{Hopkins1984}     & terrestrial &       & No    & Host-parasitoid web \\
        151   & 3 & \cite{Richards1926}     & terrestrial & 51,20,04 N & Yes   &       \\
        152   & 3 & \cite{Whittaker1984}    & terrestrial &       & No    & Source web \\
        153   & 3 & \cite{Mayse1978}    & terrestrial & 40,06,38 N & Yes   &       \\
        154   & 3 & \cite{Mayse1978}    & terrestrial & 39,56,41 N & Yes   &       \\
        155   & 3 & \cite{Askew1975}     & terrestrial &       & No    & Host-parasitoid web \\
        156   & 3 & \cite{Askew1961}     & terrestrial &       & No    & Host-parasitoid web \\
        157   & 3 & \cite{Askew1961}     & terrestrial &       & No    & Host-parasitoid web \\
        158   & 3 & \cite{Askew1961}     & terrestrial &       & No    & Host-parasitoid web \\
        159   & 3 & \cite{Askew1961}     & terrestrial &       & No    & Host-parasitoid web \\
        160   & 3 & \cite{Askew1961}     & terrestrial &       & No    & Host-parasitoid web \\
        161   & 3 & \cite{Askew1961}     & terrestrial &       & No    & Host-parasitoid web \\
        162   & 3 & \cite{Askew1961}     & terrestrial &       & No    & Host-parasitoid web \\
        163   & 3 & \cite{Askew1961}     & terrestrial &       & No    & Host-parasitoid web \\
        164   & 3 & \cite{Askew1961}     & terrestrial &       & No    & Host-parasitoid web \\
        165   & 3 & \cite{Askew1961}     & terrestrial &       & No    & Host-parasitoid web \\
        166   & 3 & \cite{Askew1961}     & terrestrial &       & No    & Host-parasitoid web \\
        167   & 3 & \cite{Askew1961}     & terrestrial &       & No    & Host-parasitoid web \\
        168   & 3 & \cite{Hawkins1984}    & terrestrial &       & No    & Host-parasitoid web \\
        169   & 3 & \cite{Hawkins1984}    & terrestrial &       & No    & Host-parasitoid web \\        
        170   & 3 & \cite{Hawkins1984}    & terrestrial &       & No    & Host-parasitoid web \\
        171   & 3 & \cite{Hawkins1984}    & terrestrial &       & No    & Host-parasitoid web \\      
        172   & 3 & \cite{Hawkins1984}    & terrestrial &       & No    & Host-parasitoid web \\
        173   & 3 & \cite{Hawkins1984}    & terrestrial &       & No    & Host-parasitoid web \\
        174   & 3 & \cite{Hawkins1984}    & terrestrial &       & No    & Host-parasitoid web \\
        175   & 3 & \cite{Hawkins1984}    & terrestrial &       & No    & Host-parasitoid web \\
        176   & 3 & \cite{Hawkins1984}    & terrestrial &       & No    & Host-parasitoid web \\
        177   & 3 & \cite{Hawkins1984}    & terrestrial &       & No    & Host-parasitoid web \\
        178   & 3 & \cite{Hawkins1984}    & terrestrial &       & No    & Host-parasitoid web \\
        \hline
        \end{tabular}}%
      \end{table}

        \newpage

    \begin{table}[h!]
    \centering
    {\footnotesize
      \begin{tabular}{p{2.8cm}p{1.3cm}p{5.5cm}p{2.2cm}p{2.5cm}lp{3.5cm}}
        \hline
        Web ID & Collection & Original Source & Ecosystem & Latitude(degrees) & Used  & Reason for rejection  \\
        \hline
        179   & 3 & \cite{Robinson1953}  & terrestrial &       & No    & Source web \\
        180   & 3 & \cite{Savely1939}    & terrestrial & 36,01,02 N & Yes   &       \\
        181   & 3 & \cite{Savely1939}    & terrestrial & 36,01,02 N & Yes   &       \\
        182   & 3 & \cite{Savely1939}    & terrestrial & 36,01,02 N & Yes   &       \\
        183   & 3 & \cite{Savely1939}    & terrestrial & 36,01,02 N & Yes   &       \\
        184   & 3 & \cite{Beaver1972}     & terrestrial &       & No    & Could not locate original source \\
        185   & 3 & \cite{Chapman1955}     & terrestrial &       & No    & Source web \\
        186   & 3 & \cite{Cornaby1974}  & terrestrial &       & No    & Source web \\
        187   & 3 & \cite{Cornaby1974}  & terrestrial &       & No    & Source web \\
        188   & 3 & \cite{Jiron1981}  & terrestrial &       & No    & Source web \\
        189   & 3 & \cite{McKinnerney1978}    & terrestrial &       & No    & Could not locate original source \\
        190   & 3 & \cite{McKinnerney1978}    & terrestrial &       & No    & Could not locate original source \\
        191   & 3 & \cite{McKinnerney1978}    & terrestrial &       & No    & Could not locate original source \\
        192   & 3 & \cite{McKinnerney1978}    & terrestrial &       & No    & Could not locate original source \\
        193   & 3 & \cite{McKinnerney1978}    & terrestrial &       & No    & Could not locate original source \\
        194   & 3 & \cite{McKinnerney1978}    & terrestrial &       & No    & Could not locate original source \\
        195   & 3 & \cite{Schoenly1983a}    & terrestrial &       & No    & Could not locate original source \\
        196   & 3 & \cite{Schoenly1983a}    & terrestrial &       & No    & Could not locate original source \\
        197   & 3 & \cite{Mohr1943}    & terrestrial &       & No    & Source web \\
        198   & 3 & \cite{Schoenly1983}  & terrestrial &       & No    & Source web \\
        199   & 3 & \cite{Valiela1969}    & terrestrial &       & No    & Source web \\
        200   & 3 & \cite{Valiela1974}    & terrestrial &       & No    & Source web \\
        201   & 3 & \cite{Valiela1974}    & terrestrial &       & No    & Source web \\
        202   & 3 & \cite{Allan1982}    & stream & 38,50,21 N & Yes   &       \\
        203   & 3 & \cite{Collins1976}    & stream & 44,24 N & Yes   &       \\
        204   & 3 & \cite{Fryer1959} & lake  & 11,36,39 S & Yes   &       \\
        205   & 3 & \cite{Hildrew1985}     & stream & 51,01,05 N & Yes   &       \\
        \hline
      \end{tabular}}%
      \end{table}

        \newpage

    \begin{table}[h!]
    \centering
    {\footnotesize
      \begin{tabular}{p{2.8cm}p{1.3cm}p{5.5cm}p{2.2cm}p{2.5cm}lp{3.5cm}}
        \hline
        Web ID & Collection & Original Source & Ecosystem & Latitude(degrees) & Used  & Reason for rejection  \\
        \hline
        206   & 3 & \cite{Jones1949}     & stream & 51,41 - 51,55 N & Yes   &       \\
        207   & 3 & \cite{Koslucher1973}    & stream & 41,57 - 42,14 N & Yes   &       \\
        208   & 3 & \cite{Minckley1963}    & stream & 37,56 N & Yes   &       \\
        209   & 3 & \cite{Minshall1967}    & stream & 37,18,26 N & Yes   &       \\
        210   & 3 & \cite{Percival1929}     & stream &       & No    & Generalised scheme \\
        211   & 3 & \cite{Ricker1935}  & stream &       & No    & Could not locate original source \\
        212   & 3 & \cite{Ricker1935}  & stream &       & No    & Could not locate original source \\
        213   & 3 & \cite{Badcock1949}     & stream & 53 N  & Yes   &       \\
        214   & 3 & \cite{Thompson2003}    & stream & 44,40,44 N & Yes   &       \\
        215   & 3 & \cite{Thompson2003}    & stream & 40,40,52 N & Yes   &       \\
        216   & 3 & \cite{Thompson2003}    & stream & 35,02,50 N & Yes   &       \\
        217   & 3 & \cite{Thompson2003}    & stream & 35,03,50 N & Yes   &       \\
        218   & 3 & \cite{Thompson2004c}  & stream & 45.79 S & Yes   &       \\
        219   & 3 & \cite{Thompson2004c}  & stream & 44.64 S & Yes   &       \\
        220   & 3 & \cite{Thompson2004c}  & stream & 46.29 S & Yes   &       \\
        221   & 3 & \cite{Thompson2004c}  & stream & 46.41 S & Yes   &       \\
        222   & 3 & \cite{Thompson2004c}  & stream & 45,57,0 S & Yes   &       \\
        223   & 3 & \cite{Thompson2004c}  & stream & 46.08 S & Yes   &       \\
        224   & 3 & \cite{Thompson2004c}  & stream & 46.08 S & Yes   &       \\
        225   & 3 & \cite{Thompson2004c}  & stream & 46.115 S & Yes   &       \\
        226   & 3 & \cite{Thompson2004c}  & stream & 45.61 S & Yes   &       \\
        227   & 3 & \cite{Thompson2004c}  & stream & 44.97 S & Yes   &       \\
        228   & 3 & \cite{Thompson2004c}  & stream & 44.95 S & Yes   &       \\
        229   & 3 & \cite{Thompson2004c}  & stream & 44.96 S & Yes   &       \\
        230   & 3 & \cite{Thompson2004c}  & stream & 45 S  & Yes   &       \\
        231   & 3 & \cite{Thompson2004c}  & stream & 45.58 S & Yes   &       \\
        232   & 3 & \cite{Thompson2004c}  & stream & 45.61 S & Yes   &       \\
        233   & 3 & \cite{Thompson2004c}  & stream & 45.782 S & Yes   &       \\
        234   & 3 & \cite{Thompson2004c}  & stream & 45.803 S & Yes   &       \\
        235   & 3 & \cite{Thompson2004c}  & stream & 45.752 S & Yes   &       \\
        236   & 3 & \cite{Thompson2004c}  & stream & 46.29 S & Yes   &       \\
        237   & 3 & \cite{Thompson2004c}  & stream & 46.41 S & Yes   &       \\
        238   & 3 & \cite{Thompson2004c}  & stream & 45,57,0 S & Yes   &       \\
        239   & 3 & \cite{Thompson2004c}  & stream & 46.08 S & Yes   &       \\
         \hline
      \end{tabular}}%
      \end{table}

        \newpage

    \begin{table}[h!]
    \centering
    {\footnotesize
      \begin{tabular}{p{2.8cm}p{1.3cm}p{5.5cm}p{2.2cm}p{2.5cm}lp{3.5cm}}
        \hline
        Web ID & Collection & Original Source & Ecosystem & Latitude(degrees) & Used  & Reason for rejection  \\
        \hline
        240   & 3 & \cite{Thompson2004c}  & stream & 46.08 S & Yes   &       \\
        241   & 3 & \cite{Thompson2004c}  & stream & 46.115 S & Yes   &       \\
        242   & 3 & \cite{Thompson2004c}  & stream & 46.29 S & Yes   &       \\
        243   & 3 & \cite{Thompson2004c}  & stream & 46.41 S & Yes   &       \\
        244   & 3 & \cite{Thompson2004c}  & stream & 45,57,0 S & Yes   &       \\
        245   & 3 & \cite{Thompson2004c}  & stream & 46.08 S & Yes   &       \\
        246   & 3 & \cite{Thompson2004c}  & stream & 46.08 S & Yes   &       \\
        247   & 3 & \cite{Thompson2004c}  & stream & 46.115 S & Yes   &       \\
        248   & 3 & \cite{Thompson2004}  & estuary & 45,51,28 S & Yes   &        \\
        249   & 3 & \cite{Angelini2005}  & stream & 24,30 S & Yes   &       \\
        250   & 3 & \cite{Angelini2006}  & lake  & 22,57,38 S & Yes   &       \\
        251   & 3 & \cite{Angelini2006}  & lake  & 22,57,38 S & Yes   &       \\
        252   & 3 & \cite{Angelini2010}  & stream & 18,40 S & Yes   &       \\
        253   & 3 & \cite{Angelini2011}  & marine & 5 - 17,15 S & Yes   &       \\
        254   & 3 & \cite{Angelini2013}  & stream & 21 S  & Yes   &       \\
        255   & 3 & \cite{Angelini2013}  & stream & 21 S  & Yes   &       \\
        256   & 3 & \cite{Angelini2013}  & stream &       & No    & Could not locate original source \\
        257   & 3 & \cite{Baeta2011}  & estuary & 40,08 N & Yes   &       \\
        258   & 3 & \cite{Baeta2011}  & estuary & 40,08 N & Yes   &       \\
        259   & 3 & \cite{Baeta2011}  & estuary & 40,08 N & Yes   &       \\
        260   & 3 & \cite{Baeta2011}  & estuary & 40,08 N & Yes   &       \\
        261   & 3 & \cite{Baeta2011}  & estuary & 40,08 N & Yes   &       \\
        262   & 3 & \cite{Baeta2011}  & estuary & 40,08 N & Yes   &       \\
        263   & 3 & \cite{Schneider1997}    & lake  & 45,58,50 N & Yes   &       \\
        264   & 3 & \cite{Stagliano2002}    & terrestrial & 39,06,27 N & Yes   &       \\
        265   & 3 & \cite{Lin2006}  & marine & 22,27 N & Yes   &       \\
        266   & 3 & \cite{Cornejo-Donoso2008}  & marine & 60 - 69 S & Yes   &       \\
        267   & 3 & \cite{Zetina-Rejon2003}  & marine & 23.15 - 22.9N & Yes   &       \\
        268   & 3 & \cite{Cruz-Escalona2007}  & marine & 18,45 - 18,50 N & Yes   &       \\
        269   & 3 & \cite{Liu2007}  & lake  & 29,32,35 N & Yes   &       \\
         \hline
      \end{tabular}}%
      \end{table}

        \newpage

    \begin{table}[h!]
    \centering
    {\footnotesize
      \begin{tabular}{p{2.8cm}p{1.3cm}p{5.5cm}p{2.2cm}p{2.5cm}lp{3.5cm}}
        \hline
        Web ID & Collection & Original Source & Ecosystem & Latitude(degrees) & Used  & Reason for rejection  \\
        \hline
        270   & 3 & \cite{Filgueira2011}  & marine & 42,17,37 - 42,19,29 N & Yes   &       \\
        271   & 3 & \cite{Amundsen2013}  & lake  & 69,07 N & Yes   &        \\
        272   & 3 & \cite{Hampton2011}    & lake  & 48,03,29 N & Yes   &       \\
        273   & 3 & \cite{Parker2006}  & stream & 69,1 N & Yes   &       \\
        274   & 3 & \cite{Parker2006}  & stream & 69,1 N & Yes   &       \\
        275   & 3 & \cite{Parker2006}  & stream & 69,1 N & Yes   &       \\
        276   & 3 & \cite{Parker2006}  & stream & 69,1 N & Yes   &       \\
        277   & 3 & \cite{Massana1996}  & lake  & 42,08,6 - 42,07 N & Yes   &       \\
        278   & 3 & \cite{Stewart2011}    & lake  & 43.7 N & Yes   &       \\
        279   & 3 & \cite{Stewart2011}    & lake  & 43.7 N & Yes   &       \\
        280   & 3 & \cite{Cromar1996}  & stream & 43,58 N & Yes   &       \\
        281   & 3 & \cite{Cromar1996}  & stream & 44,58 N & Yes   &       \\
        282   & 3 & \cite{Cromar1996}  & stream & 45,58 N & Yes   &       \\
        283   & 3 & \cite{Cromar1996}  & stream & 46,58 N & Yes   &       \\
        284   & 3 & \cite{Cromar1996}  & stream & 47,58 N & Yes   &       \\
        285   & 3 & \cite{Cromar1996}  & stream & 48,58 N & Yes   &       \\
        286   & 3 & \cite{Cromar1996}  & stream & 49,58 N & Yes   &       \\
        287   & 3 & \cite{Cromar1996}  & stream & 50,58 N & Yes   &       \\
        288   & 3 & \cite{Christian1999}    & estuary & 30,05,39 N & Yes   &       \\
        289   & 3 & \cite{Fetahi2011}  & lake  & 11,15 N & Yes   &       \\
        290   & 3 & \cite{Brodeur1992}    & marine & 43 - 48 N & Yes   &       \\
        291   & 3 & \cite{Brodeur1992}    & marine & 44 - 48 N & Yes   &       \\
        292   & 3 & \cite{Brodeur1992}    & marine & 45 - 48 N & Yes   &       \\
        293   & 3 & \cite{Brodeur1992}    & marine & 46 - 48 N & Yes   &       \\
        294   & 3 & \cite{Gaedke1994}  & lake  & 47,37,30 N & Yes   &       \\
        295   & 3 & \cite{Preston2012}    & lake  & 37.66 N & Yes   &        \\
        296   & 3 & \cite{Ratsirarson1996}  & lake  & 24,55 S & Yes   &       \\
        297   & 3 & \cite{Closs1994}  & stream & 37,40 S & Yes   &       \\
        298   & 3 & \cite{Closs1994}  & stream & 37,40 S & Yes   &       \\
        299   & 3 & \cite{Closs1994}  & stream & 37,40 S & Yes   &       \\
        300   & 3 & \cite{Closs1994}  & stream & 37,40 S & Yes   &       \\
        301   & 3 & \cite{Closs1994}  & stream & 37,40 S & Yes   &       \\
        302   & 3 & \cite{Closs1994}  & stream & 37,40 S & Yes   &       \\
        303   & 3 & \cite{Closs1994}  & stream & 37,40 S & Yes   &       \\
         \hline
      \end{tabular}}%
      \end{table}

        \newpage

    \begin{table}[h!]
    \centering
    {\footnotesize
      \begin{tabular}{p{2.8cm}p{1.3cm}p{5.5cm}p{2.2cm}p{2.5cm}lp{3.5cm}}
        \hline
        Web ID & Collection & Original Source & Ecosystem & Latitude(degrees) & Used  & Reason for rejection  \\
        \hline
        304   & 3 & \cite{Closs1994}  & stream & 37,40 S & Yes   &       \\
        305   & 3 & \cite{Closs1994}  & stream & 37,40 S & Yes   &       \\
        306   & 3 & \cite{Closs1994}  & stream & 37,40 S & Yes   &       \\
        307   & 3 & \cite{Closs1994}  & stream & 37,40 S & Yes   &       \\
        308   & 3 & \cite{Closs1994}  & stream & 37,40 S & Yes   &       \\
        309   & 3 & \cite{Gontikaki2011}  & marine & 61,13,88 N & Yes   &       \\
        310   & 3 & \cite{Khan2009}  & lake  & 12,45,25 N & Yes   &       \\
        311   & 3 & \cite{Memmott2000}     & terrestrial &       & No    & Source web \\
        312   & 3 & \cite{Kaiser-Bunbury2009}  & terrestrial &       & No    & Plant-pollinator web \\
        313   & 3 & \cite{Kaiser-Bunbury2009}  & terrestrial &       & No    & Plant-pollinator web \\
        314   & 3 & \cite{Kaiser-Bunbury2011}  & terrestrial &       & No    & Plant-pollinator web \\
        315   & 3 & \cite{Kaiser-Bunbury2011}  & terrestrial &       & No    & Plant-pollinator web \\
        316   & 3 & \cite{Kaiser-Bunbury2011}  & terrestrial &       & No    & Plant-pollinator web \\
        317   & 3 & \cite{Kaiser-Bunbury2011}  & terrestrial &       & No    & Plant-pollinator web \\
        318   & 3 & \cite{Kaiser-Bunbury2011}  & terrestrial &       & No    & Plant-pollinator web \\
        319   & 3 & \cite{Kaiser-Bunbury2011}  & terrestrial &       & No    & Plant-pollinator web \\
        320   & 3 & \cite{Ruzicka2012}    & marine & 42 - 49 N & Yes   &       \\
        321   & 3 & \cite{Ruzicka2012}    & marine & 43 - 49 N & Yes   &       \\
        322   & 3 & \cite{Ruzicka2012}    & marine & 44 - 49 N & Yes   &       \\
        323   & 3 & \cite{Ruzicka2012}    & marine & 45 - 49 N & Yes   &       \\
        324   & 3 & \cite{Ruzicka2012}    & marine & 46 - 49 N & Yes   &       \\
        325   & 3 & \cite{Lewis2002}  & terrestrial &       & No    & Host-parasitoid web \\
        326   & 3 & \cite{Lewis2002}  & terrestrial &       & No    & Host-parasitoid web \\
        327   & 3 & \cite{Kelleway2010}  & stream & 29,13 S & Yes   &       \\
        328   & 3 & \cite{Kelleway2010}  & stream & 29,20,0 S & Yes   &       \\
        329   & 3 & \cite{Kelleway2010}  & stream & 29,35 S & Yes   &       \\
        330   & 3 & \cite{Kelleway2010}  & stream & 29,13 S & Yes   &       \\
        331   & 3 & \cite{Kelleway2010}  & stream & 29,20,0 S & Yes   &       \\
        332   & 3 & \cite{Huang2008}  & stream & 31.70 N & Yes   &       \\
        333   & 3 & \cite{Link2002}    & marine & 33,51 - 44,49 N & Yes   &       \\
        334   & 3 & \cite{Alcorlo2001}  & lake  & 41,26,36 N & Yes   &       \\
        335   & 3 & \cite{Alcorlo2001}  & lake  & 41,26,05 N & Yes   &       \\
        336   & 3 & \cite{Alcorlo2001}  & lake  & 41,26,36 N & Yes   &       \\
        337   & 3 & \cite{Alcorlo2001}  & lake  & 41,26,05 N & Yes   &       \\
         \hline
      \end{tabular}}%
      \end{table}

        \newpage

    \begin{table}[h!]
    \centering
    {\footnotesize
      \begin{tabular}{p{2.8cm}p{1.3cm}p{5.5cm}p{2.2cm}p{2.5cm}lp{3.5cm}}
        \hline
        Web ID & Collection & Original Source & Ecosystem & Latitude(degrees) & Used  & Reason for rejection  \\
        \hline
        338   & 3 & \cite{Torres2013}  & marine &       & No    & Could not locate original source \\
        339   & 3 & \cite{Bulman2001}  & marine & 37-39 S & Yes   &       \\
        340   & 3 & \cite{Smith2007}  & marine & 74-78 S & Yes   &       \\
        341   & 3 & \cite{Lancraft1991}  & marine & 58,20.8 - 60,10.4 S & Yes   &       \\
        342   & 3 & \cite{Hopkins1993}  & marine & 57,00 - 60,45 S & Yes   &       \\
        343   & 3 & \cite{Kitching1987}  & lake  & 28,12 S & Yes   &       \\
        344   & 3 & \cite{Hodkinson2004}  & terrestrial &       & No    & Host-parasitoid web \\
        345   & 3 & \cite{Boit2012}  & lake  & 47,38,10 N & Yes   &       \\
        346   & 3 & \cite{Rayner2010}  & stream & 17,7 S & Yes   &       \\
        347   & 3 & \cite{Motta2005}  & stream & 23,03 S & Yes   &       \\
        348   & 3 & \cite{Douglass2011}    & marine & 37,13 N & Yes   &       \\
        349   & 3 & \cite{Douglass2011}    & marine & 37,13 N & Yes   &       \\
        350   & 3 & \cite{Douglass2011}    & marine & 37,13 N & Yes   &       \\
        351   & 3 & \cite{Warren1989}     & lake  &       & No    & Could not locate original source \\
        352   & 3 & \cite{Poepperl2003}  & stream & 54,06 N & No    & describes compartments, not species \\
        353   & 3 & \cite{Poepperl2003}  & stream & 54,06 N & No    & describes compartments, not species \\
        354   & 3 & \cite{Goldwasser1993}  & terrestrial & 18,03 N & Yes   &       \\
        355   & 3 & \cite{Cohen2003}    & lake  & 46,13 N & Yes   &       \\
        356   & 3 & \cite{Cohen2003}    & lake  & 46,13 N & Yes   &       \\
        357   & 3 & \cite{Yodzis1998} & marine & 34,21 - 16 S & No    & Study site unclear \\
        358   & 3 & \cite{Schroter2003}  & terrestrial &       & No    & Generalised scheme \\
        359   & 3 & \cite{Baiser2012}  & lake  &       & No    & Study site unclear \\
        \hline
      \end{tabular}%
        }
    \end{table}%

  \end{landscape}
\newpage

\subsection*{Appendix S2: forms of scaling relationships with species richness}

  \subsubsection*{Methods}

    The scaling relationships between link density ($Z$) and species richness ($S$)
    has been shown to be a power law~\cite{Riede2010} of the form

    \begin{equation}
    \label{Power}
    Z_{i} \sim \alpha S_{i}^{\beta}  ,
    \end{equation}

    \noindent which is often re-expressed in logarithmic form

    \begin{equation}
    \label{Loglog}
    \log{Z_{i}} \sim \log{\alpha} + \beta\log{S_{i}}  .
    \end{equation}

    \noindent Although these relationships are very similar, they imply different error distributions~\cite{Xiao2011}.
    Specifically, equation~(\ref{Power}) implies a normally-distributed, additive 
    error and equation~(\ref{Loglog}) a lognormal,
    multiplicative error. As we have no \emph{a priori} reason to believe that our dataset 
    has one error distribution over another, we follow the recommendations in~\cite{Xiao2011} and compared the two
    model formulations explicitly. The model with the error distribution that best approximates the one
    observed in the empirical
    data was then used to test for potential effects of latitude.

    Although scaling relationships between species richness and generality or
    species richness and vulnerability have not been explicitly examined (but see scaling 
    relationships for the standard deviations of each property in~\cite{Riede2010}), we expect that they will follow
    power laws similar to that of the relationship between species richness and links per species.
    This is because the links taken into account in calculating generality and vulnerability are subsets 
    of the total links included when calculating links per species. As with links per species, we explicitly 
    compared the error distributions of models for generality and vulnerability using
    both the power-law and logarithmic formulations. 
    In each case, we used the best-fitting equation as a template when assessing the effect of latitude on scaling with
    species richness.

  \subsubsection*{Results}

    When considering the relationships between species richness and all response variables 
    (link density, generality, vulnerability), equation~(\ref{Loglog}) had a
    lower AIC than did equation~(\ref{Power}). This indicates that the
    data support an assumption of multiplicative lognormal error better than an
    assumption of additive normal error. That is, models where $\epsilon$ is
    modelled as an additive term on
    the logarithmic scale provide a better description of the data than models
    where $\epsilon$ is modelled as an additive term on the arithmetic scale.  
    We therefore used logarithmic-form models when assessing the
    effect of latitude on scaling relationships  with species richness.

\newpage

\subsection*{Appendix S3: testing for effects of sampling effort}

  \subsubsection*{Methods}
    As our dataset consists of food webs published over several decades, from 
    many countries, and by many research groups, there are
    differences in the procedures used to assemble the food
    webs and likely also differences in the amount of
    sampling effort allocated to each web. In order to 
    ensure that our results were not driven by
    particularly unusual food webs, or by unusually well- or poorly-sampled webs 
    derived from any particular research group, we performed two jackknife tests of our 
    results. In the first test, we removed one web from our dataset and re-fit 
    our models, then repeated this process for all webs in the dataset. We then 
    compared the estimates for each coefficient in each of our models in order to 
    determine whether the omission of any web dramatically changed the resulting 
    model and conclusions. In the second test, we removed groups of webs assembled by a given 
    author (or authors, where two or more authors assembled identical sets of 
    food webs) and again plotted the resulting collection of estimates for each
    coefficient. We did not plot the coefficients resulting from the removal of
    authors who compiled a single food web, as these were included in the first
    test. This allowed us to determine whether the inclusion or exclusion
    of the webs assembled by a particular researcher or group of researchers
    affected our results.

  \subsubsection*{Results}

    None of our models were greatly affected by the removal of any single food web
    or by the removal of a set of webs compiled by a common author (or group of
    authors).
    Estimates of coefficients from all jackknifed datasets were extremely
    similar (Figs.~\ref{LS_web},\ref{Gen_web},\ref{Vul_web},\ref{LS_auth},\ref{Gen_auth},\ref{Vul_auth}). 
    We are therefore very confident that, while there are undoubtedly differences 
    in the methods used to compile food webs, these differences do not explain the trends we report.

  \newpage

    \begin{figure}[!h]
    \centerline{\includegraphics*[height=.75\textheight]{Figures/Jackknife/LS_web.eps}}
    \caption{Estimates of the coefficients for scaling of link density with species richness
    in equation 5 (main text) do not vary significantly when jackknifing across food webs. 
    We show the pseudovalues for each coefficient in the equation after the removal of each 
    web in the dataset. The estimate for the full, non-jackknifed dataset is indicated by 
    the red, dashed line.}
    \label{LS_web}
    \end{figure}

  \newpage

    \begin{figure}[!h]
    \centerline{\includegraphics*[height=.75\textheight]{Figures/Jackknife/Gen_web.eps}}
    \caption{Estimates of the coefficients for scaling of generality with species richness
    (see equation 5, main text) do not vary significantly
    when jackknifing across food webs. We show the pseudovalues for each coefficient
    in the equation after the removal of each web in the dataset. The estimate for the full, non-jackknifed
    dataset is indicated by the red, dashed line. }
    \label{Gen_web}
    \end{figure}

  \newpage

    \begin{figure}[!h]
    \centerline{\includegraphics*[height=.75\textheight]{Figures/Jackknife/Vul_web.eps}}
    \caption{Estimates of the coefficients for scaling of vulnerability with species richness
    (see equation 5, main text) do not vary significantly
    when jackknifing across food webs. We show the pseudovalues for each coefficient
    in the equation after the removal of each web in the dataset.  The estimate for the full, non-jackknifed 
    dataset is indicated by the red, dashed line.}
    \label{Vul_web}
    \end{figure}

  \newpage

    \begin{figure}[!h]
    \centerline{\includegraphics*[height=.75\textheight]{Figures/Jackknife/LS_author.eps}}
    \caption{Estimates of the coefficients for scaling of link density with species richness
    in equation 5 (main text) do not vary significantly
    when jackknifing across authors of the published food webs used in this study. 
    We show the pseudovalues for each coefficient
    in the equation after the removal of all webs compiled by each author (or set of authors where
    multiple authors compiled identical sets of webs) in the dataset. Authors
    who compiled only a single web were jackknifed as part of the by-web jackknife.
    The estimate for the full, non-jackknifed dataset is indicated by the red, dashed line, and
    the number of webs removed in each jackknife is given in parentheses.}
    \label{LS_auth}
    \end{figure}

  \newpage

    \begin{figure}[!h]
    \centerline{\includegraphics*[height=.75\textheight]{Figures/Jackknife/Gen_author.eps}}
    \caption{Estimates of the coefficients for scaling of generality with species richness
    (see equation 5, main text) do not vary significantly
    when jackknifing across authors of the published food webs used in this study. 
    We show the pseudovalues for each coefficient
    in the equation after the removal of all webs compiled by each author (or set of authors where
    multiple authors compiled identical sets of webs) in the dataset. Authors
    who compiled only a single web were jackknifed as part of the by-web jackknife.
    The estimate for the full, non-jackknifed dataset is indicated by the red, dashed line, and
    the number of webs removed in each jackknife is given in parentheses.}
    \label{Gen_auth}
    \end{figure}

  \newpage

    \begin{figure}[!h]
    \centerline{\includegraphics*[height=.75\textheight]{Figures/Jackknife/Vul_author.eps}}
    \caption{Estimates of the coefficients for the scaling of vulnerability with species richness
    (see equation 5, main text) do not vary significantly
    when jackknifing across authors of the published food webs used in this study. 
    We show the pseudovalues for each coefficient
    in the equation after the removal of all webs compiled by each author (or set of authors where
    multiple authors compiled identical sets of webs) in the dataset. Authors
    who compiled only a single web were jackknifed as part of the by-web jackknife.
    The estimate for the full, non-jackknifed dataset is indicated by the red, dashed line, and
    the number of webs removed in each jackknife is given in parentheses.}
    \label{Vul_auth}
    \end{figure}

% \subsection*{Appendix S4: Scaling relationships by trophic level}

%   \subsubsection*{Methods}

%     We were also interested in the ways that scaling
%     relationships with species richness might
%     be affected by changes to the distribution of species 
%     among trophic levels. [[more justification of why]]
%     To that end, we repeated our 
%     analyses replacing species richness with the
%     proportion of basal resources, proportion of 
%     intermediate consumers, or proportion of top 
%     predators. All model fitting procedures were identical 
%     to those described for species richness. The model 
%     simplicication procedure was similar to
%     the procedure described for species richness, except 
%     that in some cases there were multiple models with 
%     $\delta$AIC\textless2 that shared the minimal degrees 
%     of freedom. In these cases, we defined the 
%     best-fitting model as the model with the lowest AIC in 
%     the set of models with $\delta$AIC\textless2 and the 
%     minimal degrees of freedom.

%   \subsubsection*{Results}

%     As with species richness, specialisation scaled with
%     the numbers and proportions of species at each trophic
%     level following a lognormal distribution.
%     The relationships between latitude and the proportions of species at each
%     trophic level were more complex than the flat relationship between latitude
%     and species richness in this dataset.  The proportion of basal resources
%     increased towards the poles in stream webs
%     ($\beta_{Latitude:stream}$=0.011, $p$\textless0.001)  but did not vary
%     with latitude in estuarine, marine, lake, or terrestrial  food webs
%     ($\beta_{Latitude}$=0.001, $p$=0.080), while the proportion of
%     intermediate consumers decreased towards the poles in all ecosystem types
%     except lakes ($\beta_{Latitude}$=-0.003, $p$=0.015 and
%     $\beta_{Latitude:Lake}$=0.005, $p$=0.016). The proportion of top predators
%     in a food web,  meanwhile, decreased towards the poles in lake and  stream
%     food webs but did not vary in other ecosystem types
%     ($\beta_{Latitude}$=0.002, $p$=0.151;  $\beta_{Latitude:Lake}$=-0.006,
%     $p$=0.001; and $\beta_{Latitude:stream}$=-0.007, $p$=0.006).

%     Unlike species richness, only generality showed an overall increase with
%     increasing proportions of basal resources in a web ($\beta_0$=0.019,
%     $p$=0.859; $\beta_0$=0.383, $p$=0.001; and $\beta_0$=0.019, $p$=0.855 for
%     link density, generality, and vulnerability, respectively;
%     Fig.~\ref{TL_rescaled}). In each case, however, there were substantial
%     effects of latitude on the scaling relationships. Thus, the scaling
%     exponents for link density and vulnerability were negative for estuarine,
%     lake, and stream communities and became more negative towards the poles
%     ($\beta_{Latitude}$=-0.008, $p$\textless0.001 and $\beta_{Latitude}$=-0.008,
%     $p$\textless0.001, respectively; Fig.~\ref{B}). For the scaling of link
%     density and vulnerability in marine and terrestrial communities, the overall
%     effect of latitude on scaling was positive ($\beta_{Latitude:Marine}$=0.011,
%     $p$=0.020; $\beta_{Latitude:Terrestrial}$=0.016, $p$\textless0.001 and
%     $\beta_{Latitude:Marine}$=0.011, $p$=0.020;
%     $\beta_{Latitude:Terrestrial}$=0.016, $p$\textless0.001, respectively) such
%     that there was weak scaling at most latitudes. In contrast, the scaling
%     exponent of generality with the proportion of basal resources was positive
%     near the equator and negative near the poles in estuarine, lake, and stream
%     communities ($\beta_{Latitude}$=-0.010, $p$\textless0.001) while there was
%     very little change in the scaling exponents for marine and terrestrial
%     communities ($\beta_{Latitude:Marine}$=0.013, $p$=0.010;
%     $\beta_{Latitude:Terrestrial}$=0.019, $p$\textless0.001).

%     As with species richness, link density, generality, and vulnerability
%     generally increased with the proportion of intermediate consumers in a web
%     ($\beta_0$=0.771, $p$\textless0.001; $\beta_0$=0.349, $p$=0.235; and
%     $\beta_0$=0.771, $p$\textless0.001 respectively; Fig.~\ref{TL_rescaled}). 
%     The best-fit models for link density and vulnerability did not include
%     any effect of latitude. In contrast, the strength of the scaling relationship
%     for generality increased weakly towards the poles in estuarine, marine, and
%     terrestrial food webs ($\beta_{Latitude}$=0.008, $p$=0.204) but decreased 
%     towards the poles in lakes and streams ($\beta_{Latitude:Lake}$=-0.021, $p$=0.018
%     and $\beta_{Latitude:Stream}$=-0.035, $p$=0.001; Fig~\ref{I}).

%     In contrast to the generally positive scaling relationships above, link
%     density, generality, and vulnerability decreased as the proportion of top
%     predators increased ($\beta_0$=-0.532, $p$\textless0.001; $\beta_0$=-0.454,
%     $p$\textless0.001; and $\beta_0$=-0.532, $p$\textless0.001 respectively;
%     Fig~\ref{TL_rescaled}). For estuarine, lake, and marine food webs the
%     strength of this scaling did not vary significantly with latitude
%     ($\beta_{Latitude}$=-0.001, $p$=0.481 for link density;
%     $\beta_{Latitude}$=-0.001, $p$=0.472 for vulnerability).
%     In stream and terrestrial food webs, however, scaling was more strongly negative in the 
%     tropics and near zero near the poles ($\beta_{Latitude:Stream}$=0.012, $p$=0.001 
%     and $\beta_{Latitude:Terrestrial}$=0.009, $p$=0.015 for link density;
%     $\beta_{Latitude:Stream}$=0.012, $p$=0.001 and $\beta_{Latitude:Terrestrial}$=0.009, $p$=0.015
%     for vulnerability; Fig.~\ref{T}. The best-fit model for generality did not include any terms
%     for latitude in any ecosystem type.

\newpage

\subsection*{Appendix S5: supplemental figure}

\begin{figure}[!h]
\centerline{\includegraphics*[height=.75\textheight]{Figures/by_TL/scaling_with_S/proportions/S_fitlines_nonts_observed.eps}}
\caption{Scaling relationships for link density, generality, 
and vulnerability relative to species richness of a food web. 
For each relationship, we show observed values (white circles) and 
a simplified form of the scaling relationship described in equation 2 (\emph{Main Text}), neglecting 
any effects of habitat type or latitude (black line, N=196 food webs). See Fig. 1 (\emph{Main Text}) to compare with 
observed values correcting for the effects of habitat type and latitude. }
\label{props_v_lat_obs}
\end{figure}

% \newpage

% \begin{figure}[!h]
% \centerline{\includegraphics*[height=.5\textheight]{Figures/by_TL/scaling_with_S/proportions/TL_fitlines_nonts.eps}}
% \caption{Scaling relationships for re-scaled link density, generality, and vulnerability 
% relative to the proportions of basal resources, intermediate consumers, and top predators in a food web. 
% Link density, generality,
% and vulnerability were each re-scaled to remove the effects of latitude and ecosystem
% type. As these relationships take the form of power laws, we did this by dividing the food-web
% property (e.g. link density) by species richness raised to an exponent including the 
% effects of latitude and, where applicable, ecosystem type and the interaction between ecosystem
% type and latitude. Note that in all cases estuarine food webs were treated as the baseline 
% ecosystem type, but that at most two ecosystem types had interactions between ecosystem type and
% latitude retained in the best-fit model (see \emph{Results} for specifics). For each relationship, 
% we show the re-scaled values (white circles) as well as the overall scaling relationship using estuarine
% ecosystems as a baseline (black line, N=196 food webs). For a figure with the uncorrected values,
% see Fig.~\ref{TL_observed}.}
% \label{TL_rescaled}
% \end{figure}

% \newpage

% \begin{figure}[!h]
% \centerline{\includegraphics*[width=.8\textwidth]{Figures/by_TL/scaling_with_S/proportions/TL_fitlines_nonts_observed.eps}}
% \caption{Scaling relationships for link density, generality, 
% and vulnerability relative to the proportions of basal resources, intermediate consumers, and top predators in a food web. 
% For each relationship, we show observed values (white circles) and 
% a simplified form of the scaling relationship described in equation 2 (\emph{Main Text}), neglecting 
% any effects of habitat type or latitude (black line, N=196 food webs). See Fig.~\ref{TL_rescaled} to compare with 
% observed values correcting for the effects of habitat type and latitude. }
% \label{TL_observed}
% \end{figure}

% \newpage

% \begin{figure}[!h]
% \centerline{\includegraphics*[width=.8\textwidth]{Figures/by_TL/marginal/B_marginal_latitude_proportions.eps}}
% \caption{Changes to the scaling of link density, generality, and vulnerability across ecosystem
% types and over latitude. For each property we show the scaling exponent for the proportion of
% basal resources (black line) with its 95\% confidence interval (in grey), based on N=196 
% empirical food webs. Latitude is given in degrees
% from the equator regardless of direction.}
% \label{B}
% \end{figure}

% \begin{figure}[!h]
% \centerline{\includegraphics*[width=.8\textwidth]{Figures/by_TL/marginal/I_marginal_latitude_proportions.eps}}
% \caption{Changes to the scaling of link density, generality, and vulnerability across ecosystem
% types and over latitude. For each property we show the scaling exponent for the proportion of
% intermediate consumers (black line) with its 95\% confidence interval (in grey), based on 
% N=196 empirical food webs. Latitude is given in degrees
% from the equator regardless of direction.}
% \label{I}
% \end{figure}

% \newpage

% \begin{figure}[!h]
% \centerline{\includegraphics*[width=.8\textwidth]{Figures/by_TL/marginal/T_marginal_latitude_proportions.eps}}
% \caption{Changes to the scaling of link density, generality, and vulnerability across ecosystem
% types and over latitude. For each property we show the scaling exponent for the proportion of
% top predators (black line) with its 95\% confidence interval (in grey), based on N=196 
% empirical food webs. Latitude is given in degrees
% from the equator regardless of direction.}
% \label{T}
% \end{figure}

\newpage

\bibliographystyle{prsb}%Compile with jae.bst style file
\bibliography{noISN}% your .bib file(s)

\end{document}