### 2015-10-11 talk/ABC_2015.tex
Punched up all the figures

\documentclass{beamer}

\mode<presentation>
{
  \usetheme{Singapore}
  \renewcommand{\insertnavigation}[1]{}
  \setbeamercovered{transparent}
}

\usepackage[english]{babel}
\usepackage[absolute,overlay]{textpos}
\usepackage[latin1]{inputenc}
\usepackage{times}
\usepackage{graphicx}
\usepackage{epstopdf}
\usepackage[T1]{fontenc}
\usepackage{soul}
\usepackage{pdflscape}
\usepackage{multirow}
\usepackage{pgf}
\usepackage{tikz}

\beamertemplatenavigationsymbolsempty

\newcommand{\citations}[1]
   {
     \begin{textblock}{16}[1,1](15.75,15.25)
       \flushright{\scriptsize #1}
     \end{textblock}
   }

\definecolor{MedPurple}{RGB}{194,165,207}
\definecolor{DeepPurple}{RGB}{118,42,131}
\definecolor{LightGreen}{RGB}{166,219,160}
\definecolor{DarkGreen}{RGB}{0,68,27}
\definecolor{MedGreen}{RGB}{90,174,97}

\setbeamercolor{title}{fg=DeepPurple}
\setbeamercolor{frametitle}{fg=DeepPurple}
\setbeamercolor{structure}{fg=DarkGreen}

\setbeamercolor{*}{bg=LightGreen}

\title[Specialisation \& the Latitude-niche-breadth Hypothesis]
{Specialisation \& the Latitude-niche-breadth Hypothesis}

\author[A.R. Cirtwill, D.B. Stouffer, \& T.N. Romanuk]{\textbf{Alyssa R. Cirtwill}, Daniel B. Stouffer, \& Tamara N. Romanuk} 
\institute[]
{
 %
  Stouffer Lab\\
  School of Biological Sciences\\
  University of Canterbury\\
  Christchurch, New Zealand\\
  ~\\
  www.stoufferlab.org\\
  %
}
\date[Short Occasion] 
{23 October 2015 | ABC 2015}

\subject{Talks}

\begin{document}

% \begin{frame}
%   \titlepage
% \end{frame}

% %%Start Introduction Slides%%
\section*{Background}

  \begin{frame}{Why are there more species in the tropics?}

    \begin{center}
      \includegraphics*[width=.8\textwidth]{Figures/plant_richness.eps}

      \vspace{1cm}

    \textbf{Alyssa R. Cirtwill},\\ Daniel B. Stouffer \& Tamara N. Romanuk

    \vspace{1cm}

    23 October 2015 | ABC 2015

    \end{center}
  \end{frame}

  \begin{frame}{Why are there more species in the tropics?}

    \begin{center}
      \includegraphics*[width=.8\textwidth]{Figures/plant_richness.eps}

      \vspace{.5cm}

      {\color{white}{\Large Latitude-Niche Breadth Hypothesis}}

      \vspace{.5cm}

      Envionmental stability? 
      Productivity? 
      Speciation rates?

      \vspace{1cm}

    \end{center}
  \end{frame}

  \begin{frame}{Why are there more species in the tropics?}

    \begin{center}
      \includegraphics*[width=.8\textwidth]{Figures/plant_richness.eps}

      \vspace{.5cm}

      {\Large Latitude-Niche Breadth Hypothesis}

      \vspace{.5cm}

      {\color{purple}Envionmental stability? Productivity?}
      Speciation rates?

      \vspace{1cm}

    \end{center}
  \end{frame}

  \begin{frame}{Latitude-Niche Breadth Hypothesis}

    \begin{center}
      \includegraphics[width=.8\textwidth]{Figures/latitude_niche_breadth_1.eps}
    \end{center}

    \vspace{1.5cm}

    \tiny{Vazquez, D.P. \& Stevens, R.D. 2004. The latitudinal gradient in niche breadth: concepts and evidence. \emph{Am. Nat.}}
  \end{frame}

  \begin{frame}{Latitude-Niche Breadth Hypothesis}

    \begin{center}
      \includegraphics[width=.8\textwidth]{Figures/latitude_niche_breadth_2.eps}
    \end{center}

    \vspace{1.5cm}

    \tiny{Vazquez, D.P. \& Stevens, R.D. 2004. The latitudinal gradient in niche breadth: concepts and evidence. \emph{Am. Nat.}}
  \end{frame}

  \begin{frame}{Latitude-Niche Breadth Hypothesis}

    \begin{center}
      \includegraphics[width=.8\textwidth]{Figures/latitude_niche_breadth_3.eps}
    \end{center}

    \vspace{1.5cm}

    \tiny{Vazquez, D.P. \& Stevens, R.D. 2004. The latitudinal gradient in niche breadth: concepts and evidence. \emph{Am. Nat.}}
  \end{frame}

  \begin{frame}{Alternatively ...}

    \begin{center}
      \includegraphics[width=.8\textwidth]{Figures/latitude_niche_breadth_4.eps}
    \end{center}

    \vspace{1.5cm}

    \tiny{Davies \emph{et. al.} 2007. Productivity alters the scale dependence of the diversity-invasibility relationship. \emph{Ecology.}}
  \end{frame}

  \begin{frame}{Alternatively ...}

    \begin{center}
      \includegraphics[width=.8\textwidth]{Figures/latitude_niche_breadth_5.eps}
    \end{center}

    \vspace{1.5cm}

    \tiny{Davies \emph{et. al.} 2007. Productivity alters the scale dependence of the diversity-invasibility relationship. \emph{Ecology.}}
  \end{frame}

\section*{Food Webs}
  \begin{frame}{Does niche breadth vary with latitude?}
    \begin{columns}
    \column{.5in}
    \column{1.75in}
      \begin{center}
      \includegraphics[height=1in]{Figures/version1.eps}\\
      \vspace{.5cm}
      Greater specialisation\\in the tropics

      \vspace{.25cm}
      {\color{white}Fewer prey per species\\in the tropics}

      \end{center}
    \column{.5in}
    \column{1.75in}
      \begin{center}
      \includegraphics[height=1in]{Figures/version2.eps}\\
      \vspace{.5cm}
      Similar specialisation\\at all latitudes

      \vspace{.25cm}
      {\color{white}Equal prey per species\\at all latitudes}

      \end{center}
    \column{.5in}
    \end{columns}

  \end{frame}

  \begin{frame}{Does niche breadth vary with latitude?}
    \begin{columns}
    \column{.5in}
    \column{1.75in}
      \begin{center}
      \includegraphics[height=1in]{Figures/version1.eps}\\
      \vspace{.5cm}
      Greater specialisation\\in the tropics

      \vspace{.25cm}
      Fewer prey per species\\in the tropics

      \end{center}
    \column{.5in}
    \column{1.75in}
      \begin{center}
      \includegraphics[height=1in]{Figures/version2.eps}\\
      \vspace{.5cm}
      Similar specialisation\\at all latitudes

      \vspace{.25cm}
      Equal prey per species\\at all latitudes

      \end{center}
    \column{.5in}
    \end{columns}

  \end{frame}

  \begin{frame}{Does niche breadth vary with latitude?}

    \begin{center}
      \includegraphics*[width=.8\textwidth]{Figures/LittleRockLake.eps}

    195 food webs \\
    0-78$^\circ$ (absolute latitude) \\
    {\color{white} Mean generality (\textbf{G})}

    \end{center}

  \end{frame}

  \begin{frame}{Does niche breadth vary with latitude?}

    \begin{center}
      \includegraphics*[width=.8\textwidth]{Figures/LittleRockLake.eps}

    195 food webs \\
    0-78$^\circ$ (absolute latitude) \\
    Mean generality (\textbf{G})

    \end{center}

  \end{frame}

\section*{Scaling}
  \begin{frame}{Does niche breadth vary with latitude?}

    \begin{center}
      \includegraphics*[width=.8\textwidth]{Figures/results/Gen_vs_lat_simulated.eps}
    \end{center}

  \end{frame}

  \begin{frame}{Niche breadth varies with species richness}

    \begin{center}
      \includegraphics*[width=.654\textwidth]{Figures/results/Gen_dots_vs_S_fitline_observed.eps}

    \vspace{.3cm}
    \end{center}
      \hspace{.74cm}
      {\Large      
      $G \approx aS^{\color{purple}b}$}
  \end{frame}

  \begin{frame}{Does latitude affect scaling?}

    \begin{center}
      \includegraphics*[width=.654\textwidth]{Figures/results/Gen_dots_vs_S_fitline_observed.eps}

      \vspace{.3cm}
    \end{center}
      \hspace{.74cm}
      {\Large
      $G \approx aS^{\color{purple}b_0+b_1Latitude}$}
  \end{frame}

  \begin{frame}{Does latitude affect scaling?}
    \begin{center}
      \includegraphics*[width=.654\textwidth]{Figures/results/Gen_dots_vs_S_fitline_observed.eps}

      \vspace{.3cm}
    \end{center}
      \hspace{.74cm}
      {\Large
      $G \approx aS^{\color{purple}b_0+b_1Latitude+b_2Ecotype+b_3Latitude:Ecotype}$}

  \end{frame}

\section*{Results}
  \begin{frame}{Does latitude affect scaling?}
    \begin{center}
      \includegraphics*[width=.9\textwidth]{Figures/results/Gen_vs_S_marginal_axis.eps}

      \vspace{.8cm}
      ${\color{purple}b_0+b_1Latitude+b_2Ecotype+b_3Latitude:Ecotype}$

    \end{center}
  \end{frame}

  \begin{frame}{Does latitude affect scaling?}
    \begin{center}
      \includegraphics*[width=.9\textwidth]{Figures/results/Gen_vs_S_marginal_one.eps}

      \vspace{.8cm}
      ${\color{purple}b_0+b_1Latitude+b_2Ecotype+b_3Latitude:Ecotype}$

    \end{center}
  \end{frame}

  \begin{frame}{Does latitude affect scaling?}
    \begin{center}
      \includegraphics*[width=.9\textwidth]{Figures/results/Gen_vs_S_marginal.eps}

      \vspace{.8cm}
      ${\color{purple}b_0+b_1Latitude+b_2Ecotype+b_3Latitude:Ecotype}$

    \end{center}
  \end{frame}

\section*{Discussion}

  \begin{frame}{Estuarine, Marine, \& Terrestrial Food Webs}
    \begin{center}
      \includegraphics*[width=.9\textwidth]{Figures/results/no_effect.eps}

      \vspace{1cm}
      {\color{white} No effect of latitude on scaling\\
      Similar niche widths at all latitude\\
      \vspace{.54cm}        
      Supports Davies \emph{et. al.}, 2007:\\
      Broader niche space in the tropics}

    \end{center}
  \end{frame}

  \begin{frame}{Estuarine, Marine, \& Terrestrial Food Webs}
    \begin{center}
      \includegraphics*[width=.9\textwidth]{Figures/results/no_effect.eps}

      \vspace{1cm}
      No effect of latitude on scaling\\
      {\color{white}Similar niche widths at all latitude\\
      \vspace{.54cm}      
      Supports Davies \emph{et. al.}, 2007:\\
      Broader niche space in the tropics}

    \end{center}
  \end{frame}

  \begin{frame}{Estuarine, Marine, \& Terrestrial Food Webs}
    \begin{center}
      \includegraphics*[width=.9\textwidth]{Figures/results/no_effect.eps}

      \vspace{1cm}
      No effect of latitude on scaling\\
      Similar niche widths at all latitude\\
      \vspace{.54cm}
      {\color{white} Supports Davies \emph{et. al.}, 2007:\\
      Broader niche space in the tropics}

    \end{center}
  \end{frame}

  \begin{frame}{Estuarine, Marine, \& Terrestrial Food Webs}
    \begin{center}
      \includegraphics*[width=.9\textwidth]{Figures/results/no_effect.eps}

      \vspace{1cm}
      No effect of latitude on scaling\\
      Similar niche widths at all latitude\\
      \vspace{.54cm}      
      Supports Davies \emph{et. al.}, 2007:\\
      Broader niche space in the tropics

    \end{center}
  \end{frame}

  \begin{frame}{Lake \& Stream Food Webs}
    \begin{center}
      \includegraphics*[width=.9\textwidth]{Figures/results/effect_mod.eps}

      \vspace{1cm}
      {\color{white} Scaling increases with latitude\\
      Narrower niches in the tropics\\
      \vspace{.54cm}      
      Supports Vazquez \& Stevens 2004\\}
      {\color{white} narrower niches in the tropics}

    \end{center}
  \end{frame}

  \begin{frame}{Lake \& Stream Food Webs}
    \begin{center}
      \includegraphics*[width=.9\textwidth]{Figures/results/effect_mod.eps}

      \vspace{1cm}
      Scaling increases with latitude\\
      {\color{white} Narrower niches in the tropics\\
      \vspace{.54cm}      
      Supports Vazquez \& Stevens 2004\\}
      {\color{white} narrower niches in the tropics}

    \end{center}
  \end{frame}

  \begin{frame}{Lake \& Stream Food Webs}
    \begin{center}
      \includegraphics*[width=.9\textwidth]{Figures/results/effect_mod.eps}

      \vspace{1cm}
      Scaling increases with latitude\\
      Narrower niches in the tropics\\
      \vspace{.54cm}     
      {\color{white} Supports Vazquez \& Stevens 2004\\}
      {\color{white} narrower niches in the tropics}

    \end{center}
  \end{frame}

  \begin{frame}{Lake \& Stream Food Webs}
    \begin{center}
      \includegraphics*[width=.9\textwidth]{Figures/results/effect_mod.eps}

      \vspace{1cm}
      Scaling increases with latitude\\
      Narrower niches in the tropics\\
      \vspace{.54cm}
      Supports Vazquez \& Stevens 2004\\
      {\color{white} narrower niches in the tropics}

    \end{center}
  \end{frame}

  \begin{frame}{What's different in freshwaters?}
    Seasonality?\\
    NZ streams are weird but probably don't affect much\\
    Gape limitation?\\
  \end{frame}

\end{document}