### 2018-08-08 manuscript/erratum_letter.tex

\documentclass[12pt]{letter}

%\usepackage[britdate]{canterbury-letter}
\usepackage[britdate]{LiU-letter}
\usepackage{times}
\usepackage{letterbib}
\usepackage{geometry}
\usepackage[round]{natbib}
\usepackage{graphicx}
\geometry{a4paper}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{authblk}
\usepackage[running]{lineno}
\usepackage{amsmath,amsfonts,amssymb}
% \usepackage[margin=10pt,font=small,labelfont=bf]{caption}

%\usepackage{natbib}
% \bibpunct[; ]{(}{)}{;}{a}{,}{;}

\newenvironment{refquote}{\bigskip \begin{it}}{\end{it}\smallskip}

\newenvironment{figure}{}

\position{Postdoctoral fellow}
\department{Department of Physics, Chemistry, and Biology (IFM)}
\location{581 83 Link\"{o}ping, Sweden}
\cityzip{581 83 Link\"{o}ping, Sweden}
\telephone{}
\fax{}
\email{alyssa.cirtwill@liu.se}
\url{http://cirtwill.github.io}
\name{Alyssa R. Cirtwill}

% \position{PhD student}
% \department{School of Biological Sciences}
% \location{Private Bag 4800}
% \telephone{+64 3 364 2729}
% \fax{+64 3 364 2590}
% \email{alyssa.cirtwill@pg.canterbury.ac.nz}
% \url{http://stoufferlab.org}
% \name{Ms. Alyssa R. Cirtwill}

\newcommand{\mytitle}{\emph{Specialisation in food webs scales with species richness but not with latitude}}
\newcommand{\myjournal}{\emph{Proceedings of the Royal Society B}}

\begin{document}

\begin{letter}{\bf Professor Spencer C.H. Barrett\\
               Editor-in-Chief\\
               Proceedings of the Royal Society B\\
               6-9 Carlton House Terrace\\
               London, UK\\
               SW1Y 5AG\\
                }

\opening{Dear Prof. Barrett:}

I am writing regarding my article \emph{Latitudinal gradients in biotic niche breadth vary across ecosystem types.} (article ID \#RSPB-2015-1589.R2). The full citation details for this article are: Cirtwill, A.R., Stouffer, D.B., \& Romanuk, T.N. (2015) \emph{Latitudinal gradients in biotic niche breadth vary across ecosystem types}. \textbf{Proceedings of the Royal Society B}: 282(1819), 20151589. doi: 10.1098/rspb.2015.1589

Recently I have revisited the dataset used in this article and noticed that 75  of the webs we used (out of a sample of 263) have irregularities or errors of formatting that caused them to be incorrectly processed by our original code. Fortunately, after repeating all analyses with corrected versions of these 75 webs our results were qualitatively identical to those published and our conclusions remain unchanged. Unfortunately, the specific values given in the published text are nearly all incorrect.

I have attached a corrected version of the results and a small portion of the discussion with the original values struck through and the corrected values provided side-by-side. No further changes to the manuscript were necessary. I hope that this demonstrates that our conclusions hold and will be suitable as a corrigendum for our manuscript. I have also attached an updated supplemental information document which now includes a table (Table S2) giving details of the webs which required corrections.

My co-authors and I deeply regret not finding and correcting these errors earlier, and appreciate your kind attention in this matter.

\closing{Regards,}

\end{letter}

% \newpage

% \setcounter{page}{1}

\end{document}