### 2018-08-28 manuscript/erratum_letter.tex
going to send the letter back to Daniel

\documentclass[12pt]{letter}

%\usepackage[britdate]{canterbury-letter}
\usepackage[britdate]{LiU-letter}
\usepackage{times}
\usepackage{letterbib}
\usepackage{geometry}
\usepackage[round]{natbib}
\usepackage{graphicx}
\geometry{a4paper}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{authblk}
\usepackage[running]{lineno}
\usepackage{amsmath,amsfonts,amssymb}
% \usepackage[margin=10pt,font=small,labelfont=bf]{caption}

%\usepackage{natbib}
% \bibpunct[; ]{(}{)}{;}{a}{,}{;}

\newenvironment{refquote}{\bigskip \begin{it}}{\end{it}\smallskip}

\newenvironment{figure}{}

\position{Postdoctoral fellow}
\department{Department of Physics, Chemistry, and Biology (IFM)}
\location{581 83 Link\"{o}ping, Sweden}
\cityzip{581 83 Link\"{o}ping, Sweden}
\telephone{}
\fax{}
\email{alyssa.cirtwill@liu.se}
\url{http://cirtwill.github.io}
\name{Alyssa R. Cirtwill}

% \position{PhD student}
% \department{School of Biological Sciences}
% \location{Private Bag 4800}
% \telephone{+64 3 364 2729}
% \fax{+64 3 364 2590}
% \email{alyssa.cirtwill@pg.canterbury.ac.nz}
% \url{http://stoufferlab.org}
% \name{Ms. Alyssa R. Cirtwill}

\newcommand{\mytitle}{\emph{Specialisation in food webs scales with species richness but not with latitude}}
\newcommand{\myjournal}{\emph{Proceedings of the Royal Society B}}

\begin{document}

\begin{letter}{\bf Professor Spencer C.H. Barrett\\
               Editor-in-Chief\\
               Proceedings of the Royal Society B\\
               6-9 Carlton House Terrace\\
               London, UK\\
               SW1Y 5AG\\
                }

\opening{Dear Prof. Barrett:}

I am writing regarding my article \emph{Latitudinal gradients in biotic niche breadth vary across ecosystem types.} (article ID \#RSPB-2015-1589.R2). The full citation details for this article are: Cirtwill, A.R., Stouffer, D.B., \& Romanuk, T.N. (2015) \emph{Latitudinal gradients in biotic niche breadth vary across ecosystem types}. \textbf{Proceedings of the Royal Society B}: 282(1819), 20151589. doi: 10.1098/rspb.2015.1589

As requested, we have revised our correction by adding a section to explicitly state the purpose of the correction. In this description we refer to \emph{Table S2}, which details the food webs that were corrected and the reason for the correction. We agree with the Editor that our previous submission required a lot of effort from a reader; we had used the ''Original text'', ''Revised text'' format in response to a specific request when we first attempted to submit the correction. In the absence of any guidelines as to how corrections should be formatted, we hope that our current approach will be more user-friendly. We give the results and discussion only once, with original values presented in struck-through font and corrected values immediately following in bold. We believe that this will allow readers to see at a glance that the changes to our results are quite minor.

Thank you again for considering this correction. My co-authors and I are glad to have the opportunity to fix our manuscript and preserve the high quality of research presented in \textbf{Proceedings of the Royal Society B}.

\closing{Regards,}

\end{letter}

% \newpage

% \setcounter{page}{1}

\end{document}