### 2018-09-04 manuscript/erratum_letter.tex
resubmitted correction'

\documentclass[12pt]{letter}

%\usepackage[britdate]{canterbury-letter}
\usepackage[britdate]{LiU-letter}
\usepackage{times}
\usepackage{letterbib}
\usepackage{geometry}
\usepackage[round]{natbib}
\usepackage{graphicx}
\geometry{a4paper}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{authblk}
\usepackage[running]{lineno}
\usepackage{amsmath,amsfonts,amssymb}
% \usepackage[margin=10pt,font=small,labelfont=bf]{caption}

%\usepackage{natbib}
% \bibpunct[; ]{(}{)}{;}{a}{,}{;}

\newenvironment{refquote}{\bigskip \begin{it}}{\end{it}\smallskip}

\newenvironment{figure}{}

\position{Postdoctoral fellow}
\department{Department of Physics, Chemistry, and Biology (IFM)}
\location{581 83 Link\"{o}ping, Sweden}
\cityzip{581 83 Link\"{o}ping, Sweden}
\telephone{}
\fax{}
\email{alyssa.cirtwill@liu.se}
\url{http://cirtwill.github.io}
\name{Alyssa R. Cirtwill}

% \position{PhD student}
% \department{School of Biological Sciences}
% \location{Private Bag 4800}
% \telephone{+64 3 364 2729}
% \fax{+64 3 364 2590}
% \email{alyssa.cirtwill@pg.canterbury.ac.nz}
% \url{http://stoufferlab.org}
% \name{Ms. Alyssa R. Cirtwill}

\newcommand{\mytitle}{\emph{Specialisation in food webs scales with species richness but not with latitude}}
\newcommand{\myjournal}{\emph{Proceedings of the Royal Society B}}

\begin{document}

\begin{letter}{\bf Dr. Daniel Costa\\
               Associate Editor\\
               Proceedings of the Royal Society B\\
               6-9 Carlton House Terrace\\
               London, UK\\
               SW1Y 5AG\\
                }

\opening{Dear Dr. Costa:}

We are pleased to submit a revision to our correction to article \emph{Latitudinal gradients in biotic niche breadth vary across ecosystem types.} (article ID \#RSPB-2015-1589.R2). The full citation details for this article are: Cirtwill, A.R., Stouffer, D.B., \& Romanuk, T.N. (2015) \emph{Latitudinal gradients in biotic niche breadth vary across ecosystem types}. \textbf{Proceedings of the Royal Society B}: 282(1819), 20151589. doi: 10.1098/rspb.2015.1589

As requested, we have revised our correction by adding a section to explicitly state the purpose of the correction. In this description we refer to \emph{Table S2}, which details the food webs that were corrected and the reason for the correction. We have also restructured our correction in an effort to be more user-friendly. We agree with the Editor that our previous submission required a lot of effort from a reader; we had used the ''Original text'', ''Revised text'' format in response to a specific request when we first attempted to submit the correction. In the absence of any guidelines as to how corrections should be formatted, we now present only the corrected text (with any changes highlighted in bold). The original and corrected values for all changes are now presented in two tables which are referred to in the text. We hope that this will allow readers to see what has changed with minimal searching. We also now show the original and corrected figures side-by-side to illustrate both the minor changes to some data points and the consistency of the overall relationships reported. We believe that these changes will make it easy for any reader to see exactly what has changed and that our original conclusions hold. If the Editor has any further specific formatting changes that he would like us to implement, we are happy to do so.

Thank you again for considering this correction. My co-authors and I are glad to have the opportunity to fix our manuscript and preserve the high quality of research presented in \textbf{Proceedings of the Royal Society B}.

\closing{Regards,}

\end{letter}

% \newpage

% \setcounter{page}{1}

\end{document}